### 2020-04-22 Prep_code/nat_stories_prep.Rmd
label when words are multi-token

---
title: "R Notebook"
output: github_document
---

```{r}
library(tidyverse)
library(readr)
```
# Initial
These are used to set things up that exist in other saved files.
```{r}
#Paragraph format it

# data <- read_delim("natural_stories_words.tsv",delim="\t", col_names = c("key","value")) %>% 
#   separate(key, into=c("Story", "word_num", "col3")) %>% 
#   filter(col3=="whole") %>% 
#   select(-col3) %>% 
#   separate("value", into=c("a","b","c","d","e","f","g"), sep=" ") %>% 
#   unite(col="Word",-Story,-word_num, sep="", na.rm=T) %>% 
#   group_by(Story) %>%
#   pivot_wider(names_from=word_num,values_from=Word) %>%
#   unite(col="Text", -Story, sep=" ", na.rm=T) %>%
#   write_delim("../Materials/natural_stories_draft.tsv", delim="\t")
```

```{r words}

words <- read_delim("natural_stories_words.tsv",delim="\t", col_names = c("key","value")) %>% 
  separate(key, into=c("Story_Num", "Word_In_Story_Num", "col3"), convert=T) %>% 
  filter(col3=="whole") %>% 
  select(-col3) %>% 
  separate("value", into=c("a","b","c","d","e","f","g"), sep=" ") %>% 
  unite(col="Word",-Story_Num,-Word_In_Story_Num, sep="", na.rm=T)

nums <- 1:64%>% as.character()

labelled <- read_delim("../Materials/natural_stories_sentences.tsv", delim="\t") %>% 
  separate(Sentence, into=nums, sep=" ") %>% 
  pivot_longer(cols=`1`:`64`, names_to="Word_In_Sentence_Num",values_to="Word") %>% 
  mutate(Word_In_Sentence_Num=as.integer(Word_In_Sentence_Num)-1) %>% 
  filter(!is.na(Word)) %>% 
  group_by(Story_Num) %>% 
  mutate("Word_In_Story_Num"=row_number()) %>% 
  left_join(words,by=c("Story_Num","Word","Word_In_Story_Num"))

```

# For Maze

```{r}
# takes sentences, writes into format to do maze (also other things)
for_maze <- read_delim("../Materials/natural_stories_sentences.tsv", delim="\t") %>%
  mutate(label=str_c(Story_Num, Sentence_Num, sep="_"), item=row_number()) %>% 
  select(label, item, Sentence) %>% 
   write_delim("ns_pre_maze.txt", delim=";", col_names=F, quote_escape = F)

```

# Parsing surprisals, etc

```{r grnn}

grnn <- read_delim("grnn_out_all.txt", delim="\t", col_names=c("model_token","grnn_surp")) %>% 
  filter(model_token!="<eos>") %>% 
  mutate(num=row_number())

grnn_translator <- read_delim("ns_tokens_trans.txt", delim="\t", col_names=c("Item","Word_In_Sentence_Num", "Word","token"), col_types=c(col_character(),col_integer(), col_character(),col_character())) %>% 
  mutate(num=row_number()) %>% 
  left_join(grnn, by="num") %>% 
  mutate(grnn_surp=ifelse(model_token==token & grnn_surp>0, grnn_surp, NA)) %>% 
  select(-model_token, -num) %>% 
  separate(Item, into=c("Story_Num","Sentence_Num"), convert=T) %>% 
  left_join(labelled, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word")) %>% 
  group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
  summarize(grnn_surp=sum(grnn_surp),
            grnn_token_count=n())

```

```{r ngram}
ngram <- read_delim("ngram_out_all.txt", skip=3,delim="\t", col_names=c("sentence_no","token_no","model_token","ngram_surp")) %>% 
  mutate(num=row_number())

ngram_translator <- read_delim("ns_lower_trans.txt", delim="\t", col_names=c("Item","Word_In_Sentence_Num", "Word","token"), col_types=c(col_character(),col_integer(), col_character(),col_character())) %>% 
  mutate(num=row_number()) %>% 
  left_join(ngram, by="num") %>% 
  mutate(ngram_surp=ifelse(model_token==token & ngram_surp>0, ngram_surp, NA)) %>% 
  select(-model_token, -num, -sentence_no, -token_no) %>% 
  separate(Item, into=c("Story_Num","Sentence_Num"), convert=T) %>% 
  left_join(labelled, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word")) %>% 
  group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
  summarize(ngram_surp=sum(ngram_surp),
            ngram_token_count=n())
```

```{r txl}
txl <- read_delim("txl_out_all.txt", delim="\t", col_names=c("model_token","txl_surp")) %>% 
  filter(model_token!="<eos>") %>% 
  mutate(num=row_number())

txl_translator <- read_delim("ns_tokens_moses_trans.txt", delim="\t", 
                             col_names=c("Item","Word_In_Sentence_Num", "Word","token"), 
                             col_types=c(col_character(),col_integer(), col_character(),col_character())) %>% 
  mutate(num=row_number()) %>% 
  left_join(txl, by="num") %>% 
  mutate(txl_surp=ifelse(model_token==token & txl_surp, txl_surp, NA)) %>% 
  select(-model_token, -num) %>% 
  separate(Item, into=c("Story_Num","Sentence_Num"), convert=T) %>% 
  left_join(labelled, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word")) %>% 
  group_by(Story_Num, Sentence_Num,Word_In_Sentence_Num,Word,Word_In_Story_Num) %>% 
  summarize(txl_surp=sum(txl_surp),
            txl_token_count=n())
```

```{r freq}

freq <- read_delim("ns_freq.txt", delim="\t",
                   col_names=c("Item","Word_In_Sentence_Num","Word","freq","length"), 
                   col_types=c(col_character(),col_integer(),col_character(),col_double(), col_integer())) %>% 
  mutate(freq=ifelse(freq>0, freq, NA)) %>% 
  separate(Item, into=c("Story_Num","Sentence_Num"), convert=T) %>% 
  left_join(labelled, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word"))

```

```{r conglomerate}
all <- txl_translator %>% 
  left_join(ngram_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
  left_join(grnn_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>%
  left_join(freq, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
  filter(Word_In_Sentence_Num!=0) %>% 
  ungroup() %>% 
  mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
         freq_center=freq-mean(freq, na.rm=T),
         length_center=length-mean(length, na.rm=T)) %>% 
  write_rds("natural_stories_surprisals.rds")

```

Note that word 0's of sentences will be removed b/c we definitely don't trust them