### 2020-04-30 Analysis/nat_stories_lms.Rmd
cleaned up with only current analysis

---
title: "Natural Stories LM analysis"
output: github_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
library(mgcv)
library(mgcViz)
library(tidymv)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

```{r participants}

data <- read_rds("../Data/cleaned.rds")

data_filt <- data %>% filter(native %in% c("ENG", "English", "ENGLISH", "english")) #I peeked at what people put that semantically maps to english

data_error_summ <- data_filt %>% 
  mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
  group_by(subject, critical, practice) %>%
  filter(type!="practice") %>% 
  filter(rt<5000) %>% 
  summarize(pct_correct=mean(correct.num)) %>% 
  ungroup() %>% 
  mutate(is.attentive=ifelse(pct_correct>.8, T,F)) %>% 
  select(subject, is.attentive)

data_low_error <- data_filt %>% 
  left_join(data_error_summ, by="subject") %>% 
  filter(is.attentive) %>% 
  filter(type!="practice")

  data_error_free <- data_low_error %>% 
    mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
    group_by(sentence, subject) %>% fill(word_num_mistake) %>% ungroup() %>% 
    mutate(after_mistake=word_num-word_num_mistake,
           after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
    filter(correct=="yes") %>% 
    filter(!after_mistake %in% c(1,2))

data_no_first <- data_error_free %>% filter(word_num!=0)

data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  select(subject, word_num, word, rt, sentence, type)

data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  filter(after_mistake==0) %>% 
  select(subject, word_num, word, rt, sentence, type)

data_stories <- data_ready %>% select(type, subject) %>% 
  unique() %>% 
  group_by(type) %>% 
  tally()
```

# Overview

`r data %>% select(subject) %>% unique() %>% nrow()` participants read naturalistic stories from the natural stories corpus. Each participant read 1 story.

We exclude

- participants who do not report English as a native language (`r data_filt %>% select(subject) %>% unique() %>% nrow()` remaining)
- participants who do not get 80% of the words correct (`r data_low_error %>% select(subject) %>% unique() %>% nrow()` remaining)
- practice items (`r data_low_error %>% nrow()` words remaining)
- words that were wrong or were within two after a mistake (`r data_error_free %>% nrow()` words remaining)
- the first word of every sentence (didn't have a real distractor, RT is measured slightly differently) (`r data_no_first %>% nrow()` words remaining)
- words with RTs <100 or >5000 (<100 we think is likely a recording error, or at least not reading the words at all, >5000 is likely getting distracted) (`r data_ready %>% nrow()` words remaining)

Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times, for an average of `r mean(data_stories$n)`.

We also do the analyses on only the words before mistakes (per sentence) (`r data_pre_error %>% nrow()` words)

From the modelling side:
(After attempts without doing this filtering) we only include words which are single token and known words in each of the models vocabularies. We also only include words with frequencies. This is roughly equivalent to excluding words with punctuation.

We use as predictors:

- length in characters of stripped word
- unigram frequency of word. Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances. (This tends to tokenize off punctuation, but is capitalization sensitive). Frequencies are represented as log2 of the expected occurances in 1 billion words.

Surprisals are measured in bits.

- ngram (5-gram KN smoothed)
- GRNN
- Transformer-XL

We sometimes center the predictors, but don't rescale them. (We center them for the brm models, but don't center surprisal for the GAM model).

The model formula we use is rt ~ surp* length +  freq * length + past_surp * past_length + past_freq * past_length.

```{r labels}
labs <- read_rds("../Prep_code/natural_stories_surprisals.rds") %>% left_join(read_delim("../Materials/natural_stories_sentences.tsv", delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())

select_labs <- labs %>% 
  filter(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1) %>% 
  filter(!is.na(txl_surp) & !is.na(ngram_surp) & !is.na(grnn_surp) & !is.na(freq) & !is.na(length)) %>% 
    mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
         freq_center=freq-mean(freq, na.rm=T),
         length_center=length-mean(length, na.rm=T))

past_word <- select_labs %>% 
  mutate(word_num=word_num+1) %>% 
  rename(past_txl=txl_surp, past_ngram=ngram_surp, past_grnn=grnn_surp, past_freq=freq, past_length=length,past_c_txl=txl_center, past_c_ngram=ngram_center, past_c_grnn=grnn_center, past_c_freq=freq_center, past_c_length=length_center) %>% 
  select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
  left_join(select_labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))

labelled_pre_error <- data_pre_error %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
  select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("pre_error.rds")

labelled_post_error <- data_ready %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("post_error.rds")

```
```{r}

lab_summary <- select_labs %>% summarize_at(c("length","freq","txl_surp","grnn_surp","ngram_surp"), list(mean=mean, stdev=sd, min=min, max=max)) %>% 
  pivot_longer(everything()) %>% arrange(name)

ggplot(select_labs, aes(length))+geom_bar()

```

## BRM models

We include a by-subject effect for everything, and a by_word random intercept (full mixed effects).

Priors:

- normal(1000,1000) for intercept -- we think RTs are about 1 second usually
- normal(0,500) for beta and sd -- we don't really know what effects are
- lkj(1) for correlations -- we don't have reason to think correlations might go any particular way

```{r}
  priors <- c(
      set_prior("normal(1000, 1000)", class="Intercept"),
      set_prior("normal(0, 500)", class="b"),
      set_prior("normal(0, 500)", class="sd"),
      set_prior("lkj(1)",       class="cor"))

```

### On pre-error data only 
```{r}
brm_ngram_interact <- brm(rt ~ ngram_center* length_center + freq_center * length_center + 
                         past_c_ngram*past_c_length + past_c_freq*past_c_length + 
                         (ngram_center* length_center + freq_center * length_center + 
                            past_c_ngram *past_c_length+ past_c_freq*past_c_length|subject)+
                         (1|Word_ID), data=labelled_pre_error, 
                       file="brm_ngram_interact", prior=priors, control=list(adapt_delta=.95))

brm_grnn_interact <- brm(rt ~ grnn_center * length_center+ freq_center * length_center + 
                         past_c_grnn *past_c_length+ past_c_freq*past_c_length + 
                         (grnn_center* length_center + freq_center * length_center + 
                            past_c_grnn*past_c_length + past_c_freq*past_c_length|subject)+
                        (1|Word_ID), data=labelled_pre_error, 
                       file="brm_grnn_interact", prior=priors, control=list(adapt_delta=.95))

brm_txl_interact <- brm(rt ~ txl_center* length_center + freq_center * length_center + 
                         past_c_txl*past_c_length + past_c_freq*past_c_length + 
                         (txl_center* length_center + freq_center * length_center + 
                            past_c_txl *past_c_length+ past_c_freq*past_c_length|subject)+
                       (1|Word_ID), data=labelled_pre_error, 
                       file="brm_txl_interact", prior=priors, control=list(adapt_delta=.95))

```

```{r}

show_summary <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()

  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
    group_by(.variable) %>% 
    summarize(pct_above_0=mean(above_0)) %>% 
    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
    left_join(intervals, by=".variable") %>% 
    mutate(lower=round(.lower, digits=1),
           upper=round(.upper, digits=1),
           E=round(.value, digits=1),
           `CI`=str_c("[",lower,", ", upper,"]"),
           Term=str_sub(.variable, 3, -1),
           ) %>% 
    select(Term, `E`, `CI`,`P`)

  stats
}
a <- show_summary(brm_txl_interact) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
b <- show_summary(brm_grnn_interact) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
c <- show_summary(brm_ngram_interact) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))

summ <- a %>% union(b) %>% union(c) %>% pivot_wider(names_from="model", values_from=c(`E`,`CI`,`P`)) %>% 
  select(Term, `E_5-gram`, `CI_5-gram`, `P_5-gram`, `E_GRNN`, CI_GRNN, P_GRNN, E_TXL, CI_TXL, P_TXL)

library(xtable)
print(xtable(summ,digits=c(0,0,1,0,2,1,0,2,1,0,2)), include.rownames=FALSE)
```

### On post-error as well

```{r}
brm_ngram_interact_post <- brm(rt ~ ngram_center* length_center + freq_center * length_center + 
                         past_c_ngram*past_c_length + past_c_freq*past_c_length + 
                         (ngram_center* length_center + freq_center * length_center + 
                            past_c_ngram *past_c_length+ past_c_freq*past_c_length|subject)+
                         (1|Word_ID), data=labelled_post_error, 
                       file="brm_ngram_interact_post", prior=priors, control=list(adapt_delta=.95))

brm_grnn_interact_post <- brm(rt ~ grnn_center * length_center+ freq_center * length_center + 
                         past_c_grnn *past_c_length+ past_c_freq*past_c_length + 
                         (grnn_center* length_center + freq_center * length_center + 
                            past_c_grnn*past_c_length + past_c_freq*past_c_length|subject)+
                        (1|Word_ID), data=labelled_post_error, 
                       file="brm_grnn_interact_post", prior=priors, control=list(adapt_delta=.95))

brm_txl_interact_post <- brm(rt ~ txl_center* length_center + freq_center * length_center + 
                         past_c_txl*past_c_length + past_c_freq*past_c_length + 
                         (txl_center* length_center + freq_center * length_center + 
                            past_c_txl *past_c_length+ past_c_freq*past_c_length|subject)+
                       (1|Word_ID), data=labelled_post_error, 
                       file="brm_txl_interact_post", prior=priors, control=list(adapt_delta=.95))

```
```{r}
show_summary_2 <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()

  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
    group_by(.variable) %>% 
    summarize(pct_above_0=mean(above_0)) %>% 
    mutate(`P-value equivalent` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
    left_join(intervals, by=".variable") %>% 
    mutate(lower=round(.lower, digits=2),
           upper=round(.upper, digits=2),
           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
           Term=str_sub(.variable, 3, -1),
           Estimate=round(.value, digits=2)) %>% 
    select(Term, Estimate, `Credible Interval`, `P-value equivalent`)

  stats
}
```

### Summaries

```{r}
summary(brm_ngram_interact_post)
summary(brm_ngram_interact)

summary(brm_grnn_interact_post)
summary(brm_grnn_interact)

summary(brm_txl_interact_post)
summary(brm_txl_interact)
```

## GAMs

I think there might be ways to have mixed effects for this, but for now we're not. Also, couldn't figure out good ways of getting graphs from the full models, so just fitting the pieces within ggplot for graphs.

Using only pre-error data.

```{r gams-2}
gam_ngram_1 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)

gam_txl_1 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)

gam_grnn_1 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)

```

```{r}
for_graph <- labelled_pre_error %>% 
  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
  mutate(type=case_when(
         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
         T~"Current"),
         model=case_when(
           label %in% c("txl_surp", "past_txl")~"TXL",
           label %in% c("grnn_surp", "past_grnn")~"GRNN",
           T ~ "5-gram"
         ))

ggplot(for_graph, aes(x=data, y=rt))+
         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
         facet_grid(type~model)+
        labs(x="Surprisal (bits)", y="Reaction Time (ms)")

ggsave("../Papers/gam.pdf", width=3.5, height=2, unit="in")

```

Including post-error data.

```{r gams-3}
gam_ngram_2 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)

gam_txl_2 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)

gam_grnn_2 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)

```

```{r plot_gam-3}
for_graph_2 <- labelled_post_error %>% 
  select(rt, txl_surp, ngram_surp, grnn_surp, past_txl, past_grnn, past_ngram) %>%
  pivot_longer(cols=`txl_surp`:`past_ngram`,names_to="label", values_to="data") %>% 
  mutate(type=case_when(
         label %in% c("past_ngram", "past_txl", "past_grnn")~"Previous",
         T~"Current"),
         model=case_when(
           label %in% c("txl_surp", "past_txl")~"TXL",
           label %in% c("grnn_surp", "past_grnn")~"GRNN",
           T ~ "5-gram"
         ))

ggplot(for_graph_2, aes(x=data, y=rt))+
         geom_smooth(method="gam",formula= y~s(x, bs="cr", k=20),color="black")+
         facet_grid(type~model)+
        labs(x="Surprisal (bits)", y="Reaction Time (ms)")

```

# TODOs for future

In terms of surprisal analyses, the possible next steps are doing GAM analysis more robustly (with mixed effects, presenting results from the models) and doing DTSR.

We could also try looking at the specific syntactic features that are in the dataset (the point of natural stories was to have a bunch of idioms and weird syntax); we could maybe analyse those?