### 2020-08-26 Papers/Amlap_2020_talk/amlap_2020_talk.tex
more work on talk

\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman,t]{beamer}
\usetheme{Frankfurt}
\usecolortheme{dove}
\usepackage{appendixnumberbeamer}
%\setbeamersize{text margin left=20pt,text margin right=20pt,}

\beamertemplatenavigationsymbolsempty 
\setbeamertemplate{headline}{}
\setbeamertemplate{itemize item}{\textbullet}

\addtobeamertemplate{navigation symbols}{}{
	\ifnum\insertframenumber>\inserttotalframenumber%
	\relax
	\else%
	\usebeamerfont{footline}%
	\usebeamercolor[fg]{footline}%
	\hspace{1em}%
	\insertframenumber
	\fi%
}
\addtobeamertemplate{frametitle}{\vspace*{.5cm}}{\vspace*{.5cm}}

\setbeamercolor{footline}{fg=black}
\usepackage{soul}
\makeatletter
\let\HL\hl
\renewcommand\hl{%
	\let\set@color\beamerorig@set@color
	\let\reset@color\beamerorig@reset@color
	\HL}

\usepackage{tipa}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\mode<presentation>

  \setbeamercovered{invisible}
\usepackage{multicol}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{ulem}
\usepackage{tipa}
\usepackage{qtree}
\usepackage{phonrule}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{xcolor}
\setlength\parindent{0pt}
\usepackage{natbib}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{tcolorbox}
\tcbuselibrary{raster}

\newcommand{\bluecheck}{}%
\DeclareRobustCommand{\greencheck}{%
	\tikz\fill[scale=0.6, color=ForestGreen]
	(0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;%
}

\title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
\author{Veronica Boyce, Roger Levy}
\date{AMLaP 2020}

\begin{document}

\begin{frame}
\maketitle
%TODO maybe put a picture of Roger's face??, at least acknowledge him 
\end{frame}

\begin{frame}{Incremental processing methods}
\pause
Want a measure of processing difficulty \greencheck\pause

Assume that longer RT = more difficulty \pause
\smallskip
\begin{center}
{\large How to measure RT?}
\end{center}
\end{frame}
\begin{frame}{Incremental processing methods}

{\large Common ways to measure RT}
\begin{columns}[t]\pause
	\begin{column}{.5\textwidth}
		\begin{center}
			\textbf{\large Eye-tracking}

			\medskip

			\includegraphics[width=.9\textwidth]{../eye-tracker.jpeg}
			\pause
		\end{center}
	\end{column}\pause
	\begin{column}{.5\textwidth}
		\begin{center}
\textbf{\large Self-paced reading}

\medskip

\begin{tikzpicture}
\draw (-.5,-.5) rectangle (3.2,.5);
\onslide<4>{\node at (0,0) {The};}
\onslide<4>{\node at (1.7,0) {- - - - - - - - - - -};}
\onslide<5>{\node at (0,0) {- - - };}
\onslide<5>{\node at (.7,0) {cat};}
\onslide<5>{\node at (2,0) {- - - - - - - - };}
\onslide<6>{\node at (0.3,0) {- - - - - - };}
\onslide<6>{\node at (1.6,0) {drank};}
\onslide<6>{\node at (2.6,0) {- - - -};}
\onslide<7->{\node at (.9,0) {- - - - - - - - - - -};}
\onslide<7->{\node at (2.7,0) {milk.};}
\end{tikzpicture}
\end{center}
\end{column}
\end{columns}
\bigskip
\onslide<8->{Different methods have different trade-offs}
\end{frame}

\begin{frame}[c]{An alternative: Maze}

\pause 
\centering

\bigskip

\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}
\begin{tikzpicture}
\onslide<1-2>{\node  (c) at (-1,.5) {The};}
\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
\onslide<3-4>{\node  (d) at (1,.5) {dog};}
\onslide<5-6>{\node  (c) at (-1,.5) {revise};}
\onslide<5-6>{\node  (d) at (1,.5) {chased};}
\onslide<7-8>{\node  (c) at (-1,.5) {the};}
\onslide<7-8>{\node  (d) at (1,.5) {wish};}
\onslide<9-10>{\node  (c) at (-1,.5) {mitigate.};}
\onslide<9-10>{\node  (d) at (1,.5) {squirrel.};}
\onslide<2,8>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
\onslide<4,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
align=center] at (d) {};}
\end{tikzpicture}
\end{frame}

\begin{frame}{An alternative: Maze}
\vskip -1em
{\small(Forster et al. 2009; Witzel et al. 2012)}
	\vskip -1em
\begin{columns}
	\begin{column}{0.5\textwidth}
		\begin{center}
		\textbf{\large G-maze}\\
		`Grammatical' choices\\
		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../gmaze.pdf}
		\end{center}
	\pause
	\end{column}
	\begin{column}{0.5\textwidth} 
		\begin{center}
		\textbf{\large L-maze}\\
		`Lexical' choices\\
			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../lmaze.pdf}
		\end{center}
	\end{column}
\end{columns}

\medskip
\pause
Sentence ends if a mistake is made.
\pause

Claim: forces incremental processing (no spillover)
% Compelling results especially for G-maze, but 
% was run in the lab, and generating G-maze materials is a huge amount of work
\end{frame}

\begin{frame}{Maze Made Easy}

	Want to use Maze instead of web SPR\pause

	\medskip

	Needs some tweaks:\pause
	\begin{itemize}
		\item Run on web \pause
		\item Easily generate distractors \pause
		\item Work for multi-sentence items \pause
	\end{itemize}

\end{frame}

\begin{frame}{Run on web}
	Wrote an Ibex module \pause

	\begin{center}\includegraphics[width=.7\textwidth]{screenshot.png} \end{center}\pause

 Replicated Witzel et al. (2012) results (Boyce et al. 2020)
\end{frame}
%
\begin{frame}{Maze Made Easy}

	Want to use Maze instead of web SPR

	\medskip

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors
		\item Work for multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
	\end{itemize}

\end{frame}
%
\begin{frame}{Generating distractors}
\pause
Goal: Find a word that can't continue a partial sentence
\begin{itemize}
	\item Ex.  \textit{The dog chased} \pause
	\item Tedious (and hard!) to do by hand
\end{itemize} \pause

What makes something an unacceptable continuation? \pause
\begin{itemize}
	\item Ungrammatical \pause
	\item ...or otherwise really unlikely \pause 
	\item $\approx$ high surprisal 
\end{itemize} \pause

Can we use Neural Language Models?
\end{frame}
%
\begin{frame}{Can we use LMs?}
	\pause
Language models (LMs)
\begin{itemize}
	\item Trained to predict the next word
	\item Given a partial sentence, return probabilities of the next word
\end{itemize}\pause

LSTM RNNs show some syntactic knowledge %TODO cite people 
\pause

Add some restrictions:
\begin{itemize}
\item Restrict to a list of possible distractors
\item Only consider length, frequency matches
\end{itemize}

\end{frame}
%
\begin{frame}{Does it work?} 
\pause
\vskip -2em
{\large Yes, at least well enough.} \pause
\begin{itemize}
	\item Sometimes generates plausible distractors. \pause
	\item A-maze results comparable with G-maze {\small(Boyce et al 2020, Sloggett et al 2020)} \pause 
\end{itemize} 
\medskip
\centering

\includegraphics[width=.7\textwidth]{../a_all.pdf}
\end{frame}
%
\begin{frame}{Maze Made Easy}

	Want to use Maze instead of web SPR

	\medskip

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors \greencheck
		\item Work for multi-sentence items 
	\end{itemize}

\end{frame}
%
%
\begin{frame}{Long items}

Want to run multi-sentence items. \pause

Problem: Errors terminate sentences. \pause
\begin{itemize}
	\item Treat whole story as a unit: ~ No one makes it to the end. \pause
	\item Treat each sentence as a unit: Some participants miss key context. \pause
\end{itemize}

What if after an error, participants corrected errors and the sentence continued?

\end{frame}
%
\begin{frame}[c]{Maze with Error Correction}
\centering
\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}

\begin{tikzpicture}
\onslide<1-2>{\node  (c) at (-1,.5) {The};}
\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
\onslide<3-4>{\node  (d) at (1,.5) {dog};}
\onslide<5-8>{\node  (c) at (-1,.5) {revise};}
\onslide<5-8>{\node  (d) at (1,.5) {chased};}
\onslide<9-10>{\node  (c) at (-1,.5) {the};}
\onslide<9-10>{\node  (d) at (1,.5) {wish};}
\onslide<11-12>{\node  (c) at (-1,.5) {mitigate.};}
\onslide<11-12>{\node  (d) at (1,.5) {squirrel.};}
\onslide<2,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
\onslide<4,8,12>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
	align=center] at (d) {};}
\onslide<7,8>{\node[text=red] at (0,-1.5) {\small Incorrect. Please try again.};}
\end{tikzpicture}

\end{frame}

\begin{frame}{Maze with Error Correction}

\begin{columns}
	\begin{column}{0.4\textwidth}
		\begin{center}
		\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{../maze_diagram.pdf}
		\end{center}

	\end{column} \pause 
	\begin{column}{0.6\textwidth} 
		\begin{center}
			\begin{itemize}
				\item Can be toggled in Ibex \pause
				\item Long materials feasible \pause
				\item Have all the data \pause
				\item Compensates for bad distractors
			\end{itemize}
		\end{center}
	\end{column}
\end{columns}
\medskip
\pause
\end{frame}
%

\begin{frame}{Maze Made Easy}

	Want to use Maze instead of web SPR

	\medskip

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors \greencheck
		\item Work for multi-sentence items \greencheck ? 
	\end{itemize}

\end{frame}
%%So now that I've explained the method, going to switch over to current experiment & what we hoped to get from it
%
\begin{frame}{Current experiment}
	Various open questions to address \pause
	\begin{itemize}
		\item Are people willing to read long texts in this way? \pause
		\item Do people comprehend what they read? \pause
		\item Does the error correction variant work? \pause
		\item Do we get predictability effects? 
	\end{itemize}
\end{frame}

\begin{frame}{Natural Stories}
Natural stories corpus (Futrell et al. 2017) \pause
\begin{itemize}
	\item 10 stories, each about 1000 words \pause
	\item Some unusual constructions, but read fluently \pause
	\item 6 comprehension questions per story \pause
\end{itemize}

\end{frame}

\begin{frame}{Natural Stories}

\begin{small}Tulip mania was a period in the Dutch Golden Age during which contract prices for bulbs of the recently introduced tulip reached extraordinarily high levels and then suddenly collapsed. At the peak of tulip mania in February sixteen thirty-seven, tulip contracts sold for more than ten times the annual income of a skilled craftsman. It is generally considered the first recorded economic bubble. [...]
\medskip

Q: When did tulip mania reach its peak?

A: \hspace{3em} 1630's\hspace{3em} 1730's \end{small}

%HARD mode

\end{frame}

\begin{frame}{Participant accuracy}
100 participants from MTurk each read 1 story (20 minutes) \pause
\begin{center}
\includegraphics[width=.8\textwidth]{../error.pdf}
\end{center}
\end{frame}

\begin{frame}{Comprehension questions}
\begin{center} \pause
\includegraphics[width=.8\textwidth]{../comp.pdf}
\end{center}
% exclusion options: use comprehension checks as participant filter. Here we can use overall accuracy as the "paying attention" filter. Don't have to worry about how to write good comprehension questions. May still need to do RT exclusions for double-presses or getting distracted. 
\end{frame}

%HERE%
\begin{frame}{Surprisal Effects}
%TODO copy from updated abstract
Use 3 LMs to estimate surprisal: smoothed 5-gram, Gulordava RNN, Transformer-XL (Gulordava et al. 2018, Dai et al. 2019)
\pause

%TODO make sure that updated correct gams are used!!!

Fit GAMs on pre-error data.
\begin{itemize}
	\item Limit to single-token words.
	\item Fit to both current and past word surprisal.
\end{itemize}

\end{frame}

\begin{frame}
	Implications for surprisal theory -- size of the effects is different
	include effect sizes
	Is it about participant quality?
\end{frame}

\begin{frame}{Surprisal Effects}
\includegraphics[width=.9\textwidth]{../gam.pdf}	
\end{frame}
\begin{frame}{What about post-mistake data?}
Exclude data from mistakes or the two words after a mistake.

\includegraphics[width=.9\textwidth]{../gam2.pdf}	
\end{frame}

\begin{frame}{Summary}
\begin{itemize}\pause
	\item People will read in the Maze task for 15-20 minutes \pause
	\item It's possible to comprehend during Maze \pause
	\item Distractors are generally good enough \pause
	\item Find expected RT patterns \pause
	\item Very little spillover
\end{itemize}
\end{frame}

\begin{frame}{Use A-maze!}
Easy to use! \pause
\begin{itemize}
	\item Runs on command line \pause
	\item Match distractors across minimal pair sentences \pause
	\item Customize surprisal thresholds, vocabulary lists \pause
	\item Can output pre-formatted for Ibex
\end{itemize}
\medskip
\end{frame}

\begin{frame}{Use A-maze!} \pause
Contribute to A-maze:
\begin{itemize}
	\item Written in Python 3 \pause
	\item Interface with other language models \pause
	\item Add more frequency sources \pause
	\item Extend to non-English languages
\end{itemize}

\end{frame}

\begin{frame}{}

\textcolor{ForestGreen}{\large Documentation: vboyce.github.io/Maze}

with links to the following:
\begin{itemize}

\item A-maze code: github.com/vboyce/Maze

\item Web-maze code: github.com/vboyce/Ibex-with-Maze

\item Sample task: syntaxgym.org:666

\item Paper: psyarxiv.com/b7nqd/
\end{itemize}
\end{frame}

\appendix

\begin{frame}{Matching distractors}
If unspecified: Match by position
\begin{itemize}
	\item The son of the lady who politely introduced \sethlcolor{green} \hl{herself}\sethlcolor{pink} / \hl{himself} was popular at the party.
\end{itemize}
Can specify labels for each word to pair (within item)
\begin{itemize}
	\item The cat who the dog scared hid in a box.\\pre-1 pre-2 who art noun verb main-verb post-1 post-2 post-3
\item The dog who scared the cat sniffed around the couch.\\ pre-1 pre-2 who verb art noun main-verb post-1 post-2 post-3
\end{itemize}
\end{frame}
%%back pocket on more on model results??
%%do we even need back pockets??

\begin{frame}{Post-mistake only}
Exclude data from mistakes or the two words after a mistake.

\includegraphics[]{../gam3.pdf}	
\end{frame}

\begin{frame}{Regression coefficients}
\begin{tiny}
\begin{tabular}{l|rlr|rlr|rlr}
	\hline
	&\multicolumn{3}{c|}{5-gram}&\multicolumn{3}{c|}{GRNN}&\multicolumn{3}{c}{TXL}\\
	& Est & CI & $p$ & Est & CI & $p$ &Est & CI & $p$ \\ 
	\hline
	Intercept & 865.3 & [829.9, 902.9] & 0.00 & 871.1 & [837.9, 905.3] & 0.00 & 870.8 & [832.5, 907.8] & 0.00 \\ 
	Surprisal & 11.7 & [9.3, 14.1] & 0.00 & 23.7 & [21, 26.5] & 0.00 & 18.5 & [16.1, 21.1] & 0.00 \\ 
	Frequency & -2.9 & [-6.3, 0.5] & 0.10 & 2.9 & [-0.2, 6] & 0.06 & 0.4 & [-2.7, 3.5] & 0.79 \\ 
	Length & 20.5 & [15.4, 25.6] & 0.00 & 18.5 & [13.3, 23.7] & 0.00 & 21.4 & [16.2, 26.6] & 0.00 \\ 
	Surprisal:Length & -2.0 & [-3, -1] & 0.00 & -1.8 & [-2.7, -0.9] & 0.00 & -1.4 & [-2.2, -0.6] & 0.00 \\
	Freq:Length & -1.0 & [-2.5, 0.4] & 0.16 & -0.1 & [-1.2, 1] & 0.82 & 0.2 & [-0.9, 1.2] & 0.76 \\ 
	\hline
	Past Surprisal & 1.6 & [-0.5, 3.6] & 0.14 & 2.7 & [0.8, 4.5] & 0.00 & 0.8 & [-0.9, 2.5] & 0.40 \\ 
	Past Freq & 2.6 & [-0.1, 5.4] & 0.06 & 1.9 & [-0.2, 4.2] & 0.08 & 1.2 & [-1.1, 3.6] & 0.30 \\ 
	Past Length & -4.8 & [-9, -0.1] & 0.04 & -6.6 & [-10.9, -2.1] & 0.00 & -5.2 & [-9.3, -0.7] & 0.03 \\ 
	Past Surp:Length & -0.2 & [-1.2, 0.8] & 0.72 & -0.9 & [-1.7, -0.2] & 0.01 & -0.6 & [-1.3, 0.2] & 0.13 \\ 
	Past Freq:Length & -1.0 & [-2.3, 0.3] & 0.15 & -1.8 & [-2.9, -0.8] & 0.00 & -1.5 & [-2.6, -0.5] & 0.01 \\

	\hline
\end{tabular}
\end{tiny}
\end{frame}

\end{document}