### 2020-08-27 Papers/Amlap_2020_talk/amlap_2020_talk.tex
work on talk

\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman,t]{beamer}
\usetheme{Frankfurt}
\usecolortheme{dove}
\usepackage{appendixnumberbeamer}
%\setbeamersize{text margin left=20pt,text margin right=20pt,}

\beamertemplatenavigationsymbolsempty 
\setbeamertemplate{headline}{}
\setbeamertemplate{itemize item}{\textbullet}

\addtobeamertemplate{navigation symbols}{}{
	\ifnum\insertframenumber>\inserttotalframenumber%
	\relax
	\else%
	\usebeamerfont{footline}%
	\usebeamercolor[fg]{footline}%
	\hspace{1em}%
	\insertframenumber
	\fi%
}
\addtobeamertemplate{frametitle}{\vspace*{.5cm}}{\vspace*{.5cm}}

\setbeamercolor{footline}{fg=black}
\usepackage{soul}
\makeatletter
\let\HL\hl
\renewcommand\hl{%
	\let\set@color\beamerorig@set@color
	\let\reset@color\beamerorig@reset@color
	\HL}

\usepackage{tipa}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\mode<presentation>

  \setbeamercovered{invisible}
\usepackage{multicol}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{ulem}
\usepackage{tipa}
\usepackage{qtree}
\usepackage{phonrule}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{xcolor}
\setlength\parindent{0pt}
\usepackage{natbib}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{tcolorbox}
\tcbuselibrary{raster}

\newcommand{\bluecheck}{}%
\DeclareRobustCommand{\greencheck}{%
	\tikz\fill[scale=0.6, color=ForestGreen]
	(0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;%
}

\title{A-maze of Natural Stories: Texts are comprehensible using the Maze task}
\author{Veronica Boyce, Roger Levy}
\date{AMLaP 2020}

\begin{document}

\begin{frame}{ }
\vskip 6em

\begin{center}
{\LARGE A-maze of Natural Stories:}

\smallskip

{\Large Texts are comprehensible using\\ \smallskip the Maze task}

\bigskip

{\large Veronica Boyce, Roger Levy}

\bigskip

{\large AMLaP 2020}

\begin{tikzpicture}[remember picture,overlay]
\node[xshift=-1.8cm,yshift=-1cm] at (current page.east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{roger.jpg}};
\end{tikzpicture}

\begin{tikzpicture}[remember picture,overlay]
\node[xshift=2.5cm,yshift=-1.5cm] at (current page.north west) {\includegraphics[width=.3\textwidth]{cpl_header.png}};
\end{tikzpicture}

%TODO maybe put a picture of Roger's face??, at least acknowledge him 
\end{center}
\end{frame}

\begin{frame}{Incremental processing methods}
\pause
Want a measure of processing difficulty\pause

Assume that longer RT = more difficulty \pause
\smallskip
\begin{center}
{\large How to measure RT?}
\end{center}
\end{frame}
\begin{frame}{Incremental processing methods}

{\large Common ways to measure RT}
\begin{columns}[t]\pause
	\begin{column}{.5\textwidth}
		\begin{center}
			\textbf{\large Eye-tracking}

			\medskip

			\includegraphics[width=.9\textwidth]{../eye-tracker.jpeg}
			\pause
		\end{center}
	\end{column}
	\begin{column}{.5\textwidth}
		\begin{center}
\textbf{\large Self-paced reading}

\medskip

\begin{tikzpicture}
\draw (-.5,-.5) rectangle (3.2,.5);
\onslide<4>{\node at (0,0) {The};}
\onslide<4>{\node at (1.7,0) {- - - - - - - - - - -};}
\onslide<5>{\node at (0,0) {- - - };}
\onslide<5>{\node at (.7,0) {cat};}
\onslide<5>{\node at (2,0) {- - - - - - - - };}
\onslide<6>{\node at (0.3,0) {- - - - - - };}
\onslide<6>{\node at (1.6,0) {drank};}
\onslide<6>{\node at (2.6,0) {- - - -};}
\onslide<7->{\node at (.9,0) {- - - - - - - - - - -};}
\onslide<7->{\node at (2.7,0) {milk.};}
\end{tikzpicture}
\end{center}
\end{column}
\end{columns}
\bigskip
\onslide<8->{Different methods have different trade-offs}
\end{frame}

\begin{frame}[c]{An alternative: Maze}

\centering

\bigskip

\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}
\begin{tikzpicture}
\onslide<1-2>{\node  (c) at (-1,.5) {The};}
\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
\onslide<3-4>{\node  (d) at (1,.5) {dog};}
\onslide<5-6>{\node  (c) at (-1,.5) {revise};}
\onslide<5-6>{\node  (d) at (1,.5) {chased};}
\onslide<7-8>{\node  (c) at (-1,.5) {the};}
\onslide<7-8>{\node  (d) at (1,.5) {wish};}
\onslide<9-10>{\node  (c) at (-1,.5) {mitigate.};}
\onslide<9-10>{\node  (d) at (1,.5) {squirrel.};}
\onslide<2,8>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
\onslide<4,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
align=center] at (d) {};}
\end{tikzpicture}
\end{frame}

\begin{frame}{An alternative: Maze}
\vskip -1em
{\small(Forster et al. 2009; Witzel et al. 2012)}
	\vskip -1em
\begin{columns}
	\begin{column}{0.5\textwidth}
		\begin{center}
		\textbf{\large G-maze}\\
		`Grammatical' choices\\
		\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../gmaze.pdf}
		\end{center}
	\pause
	\end{column}
	\begin{column}{0.5\textwidth} 
		\begin{center}
		\textbf{\large L-maze}\\
		`Lexical' choices\\
			\includegraphics[clip, trim=9cm 14cm 11cm 1cm,width=.9\textwidth]{../lmaze.pdf}
		\end{center}
	\end{column}
\end{columns}

\medskip
\pause
Sentence ends if a mistake is made.
\pause

Claim: forces incremental processing (no spillover)
% Compelling results especially for G-maze, but 
% was run in the lab, and generating G-maze materials is a huge amount of work
\end{frame}

\begin{frame}{Maze Made Easy}

	Can we use Maze instead of web SPR?\pause

	\medskip

	Needs some tweaks:\pause
	\begin{itemize}
		\item Run on web \pause
		\item Easily generate distractors \pause
		\item Work for multi-sentence items
	\end{itemize}

\end{frame}

\begin{frame}{Run on web}

	\pause
	Wrote an Ibex module \pause

	\begin{center}\includegraphics[width=.7\textwidth]{screenshot.png} \end{center}\pause

 Replicated Witzel et al. (2012) results (Boyce et al. 2020)
\end{frame}
%
\begin{frame}{Maze Made Easy}

	Can we use Maze instead of web SPR?

	\medskip

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors
		\item Work for multi-sentence items %this part incidentally makes it easy to do exclusions, and might help with data loss, but we'll talk about that later
	\end{itemize}

\end{frame}
%
\begin{frame}{Generating distractors}
\pause
Goal: Find a word that can't continue a partial sentence
\begin{itemize}
	\item Ex.  \textit{The dog chased} \pause
	\item Tedious (and hard!) to do by hand
\end{itemize} \pause

What makes something an unacceptable continuation? \pause
\begin{itemize}
	\item Ungrammatical \pause
	\item ...or otherwise really unlikely \pause 
	\item $\approx$ high surprisal 
\end{itemize} \pause

Can we use Neural Language Models?
\end{frame}
%
\begin{frame}{Can we use LMs?}
	\pause
Language models (LMs)
\begin{itemize}
	\item Trained to predict the next word
	\item Given a partial sentence, return probabilities of the next word
\end{itemize}\pause

Run items through LM, choose high surprisal words as distractors
\pause

Add some restrictions:
\begin{itemize}
\item Restrict to a list of possible distractors
\item Only consider length, frequency matches
\end{itemize}

\end{frame}
%
\begin{frame}{Does it work?} 
\pause
\vskip -2em
{\large Yes, at least well enough.} \pause
\begin{itemize}
	\item Sometimes generates plausible distractors. \pause
	\item A-maze results comparable with G-maze {\small(Boyce et al 2020, Sloggett et al 2020)} \pause 
\end{itemize} 
\medskip
\centering

\includegraphics[width=.7\textwidth]{../a_all.pdf}
\end{frame}
%
\begin{frame}{Maze Made Easy}

	Can we use Maze instead of web SPR?

	\medskip

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors \greencheck
		\item Work for multi-sentence items 
	\end{itemize}

\end{frame}
%
%
\begin{frame}{Long items}

Want to run multi-sentence items. \pause

Problem: Errors terminate sentences. \pause
\begin{itemize}
	\item Treat whole story as a unit: \pause ~ No one makes it to the end. \pause
	\item Treat each sentence as a unit: \pause Some participants miss key context. \pause
\end{itemize}

What if after an error, participants corrected errors and the sentence continued?

\end{frame}
%
\begin{frame}[c]{Maze with Error Correction}
\centering
\tikzset{
	font={\fontsize{14pt}{12}\selectfont}}

\begin{tikzpicture}
\onslide<1-2>{\node  (c) at (-1,.5) {The};}
\onslide<1-2>{\node  (d) at (1,.5) {x-x-x};}
\onslide<3-4>{\node  (c) at (-1,.5) {upon};}
\onslide<3-4>{\node  (d) at (1,.5) {dog};}
\onslide<5-8>{\node  (c) at (-1,.5) {revise};}
\onslide<5-8>{\node  (d) at (1,.5) {chased};}
\onslide<9-10>{\node  (c) at (-1,.5) {the};}
\onslide<9-10>{\node  (d) at (1,.5) {wish};}
\onslide<11-12>{\node  (c) at (-1,.5) {mitigate.};}
\onslide<11-12>{\node  (d) at (1,.5) {squirrel.};}
\onslide<2,6,10>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,align=center] at (c) {};}
\onslide<4,8,12>{\node[ultra thick, draw=blue, ellipse, minimum width=60pt, minimum height= 25pt,
	align=center] at (d) {};}
\onslide<7,8>{\node[text=red] at (0,-1.5) {\small Incorrect. Please try again.};}
\end{tikzpicture}

\end{frame}

\begin{frame}{Maze with Error Correction}

\begin{columns}
	\begin{column}{0.4\textwidth}
		\begin{center}
		\includegraphics[clip, trim=9cm 12.5cm 10cm 1cm,width=\textwidth]{../maze_diagram.pdf}
		\end{center}

	\end{column} \pause 
	\begin{column}{0.6\textwidth} 
		\begin{center}
			\begin{itemize}
				\item Can be toggled in Ibex Maze \pause
				\item Long materials feasible \pause
				\item Have all the data \pause
				\item Compensates for bad distractors
			\end{itemize}
		\end{center}
	\end{column}
\end{columns}
\end{frame}
%

\begin{frame}{Maze Made Easy}

	Can we use Maze instead of web SPR?

	\medskip

	Needs some tweaks:
	\begin{itemize}
		\item Run on web \greencheck
		\item Easily generate distractors \greencheck
		\item Work for multi-sentence items \greencheck ? 
	\end{itemize}

\end{frame}
%%So now that I've explained the method, going to switch over to current experiment & what we hoped to get from it
%
\begin{frame}{Current experiment}
	Various open questions to address \pause
	\begin{itemize}
		\item Are people willing to read long texts in this way? \pause
		\item Do people comprehend what they read? \pause
		\item Does the error correction variant work? \pause
		\item Do we get predictability effects? 
	\end{itemize}
\end{frame}

\begin{frame}{Natural Stories}
Natural stories corpus (Futrell et al. 2017) \pause
\begin{itemize}
	\item 10 stories, each about 1000 words \pause
%	\item Some unusual constructions, but read fluently \pause
	\item 6 comprehension questions per story
\end{itemize}

\end{frame}

\begin{frame}{Natural Stories}

\begin{small}Tulip mania was a period in the Dutch Golden Age during which contract prices for bulbs of the recently introduced tulip reached extraordinarily high levels and then suddenly collapsed. At the peak of tulip mania in February sixteen thirty-seven, tulip contracts sold for more than ten times the annual income of a skilled craftsman. It is generally considered the first recorded economic bubble. [...]
\medskip

\pause

Q: When did tulip mania reach its peak?

A: \hspace{3em} 1630's\hspace{3em} 1730's \end{small}

%HARD mode

\end{frame}

\begin{frame}{Participant accuracy}
100 participants from MTurk each read 1 story (20 minutes) \pause
\begin{center}
\includegraphics[width=.8\textwidth]{../error.pdf}
\end{center}
\end{frame}

\begin{frame}{Comprehension questions}
\begin{center} \pause
\includegraphics[width=.8\textwidth]{../comp.pdf}
\end{center}
% exclusion options: use comprehension checks as participant filter. Here we can use overall accuracy as the "paying attention" filter. Don't have to worry about how to write good comprehension questions. May still need to do RT exclusions for double-presses or getting distracted. 
\end{frame}

%HERE%
\begin{frame}{Surprisal Effects}
\large{Is RT linear in terms of surprisal?}\pause
\medskip

Estimate surprisal from 3 models:
\begin{itemize}
	\item smoothed 5-gram
	\item LSTM-RNN (Gulordava et al. 2018)
	\item Transformer-XL (Dai et al. 2019)
\end{itemize}

\pause

Fit GAMs
\begin{itemize}
	\item Limit to single-token words
	\item Fit to both current and past word surprisal
	\item Include frequency, length as predictors
\end{itemize}

\end{frame}

\begin{frame}{Surprisal Effects}
	\includegraphics[width=.9\textwidth]{../gam.pdf}	
\end{frame}

\begin{frame}{Surprisal Effects}
	Linear Models \\
	\pause

	\begin{small}
		\begin{tabular}{l|r|r|r}
			\hline
			&5-gram & GRNN & TXL\\
			\hline
			Intercept & \textbf{865.3} & \textbf{871.1} & \textbf{870.8} \\ 
			Surprisal & \textbf{11.7}  & \textbf{23.7} & \textbf{18.5}  \\ 
			Frequency & -2.9  & 2.9  & 0.4  \\ 
			Length & \textbf{20.5} & \textbf{18.5}  & \textbf{21.4} \\ 
			Surprisal:Length & \textbf{-2.0}  & \textbf{-1.8}  & \textbf{-1.4} \\
			Freq:Length & -1.0  & -0.1 & 0.2  \\ 
			\hline
			Past Surprisal & 1.6  & \textbf{2.7} & 0.8 \\ 
			Past Freq & 2.6  & 1.9  & 1.2  \\ 
			Past Length & \textbf{-4.8}  & \textbf{-6.6 }&\textbf{ -5.2} \\ 
			Past Surp:Length & -0.2 & \textbf{-0.9} & -0.6  \\ 
			Past Freq:Length & -1.0  & \textbf{-1.8 }&   \textbf{-1.5}  \\ 
			\hline
		\end{tabular}

	Surprisal in bits, Length in characters, \\
	Frequency in $log_2$ occurrences/billion words
	\end{small}

\end{frame}
\begin{frame}{Surprisal Effects}
	Takeaways: \pause
	\begin{itemize}
		\item Minimal frequency effects (consistent with Shain 2019) \pause
		\item Long RTs \pause
		\item Large effects of Length, Surprisal \pause
		\item Very little spillover \pause
	\end{itemize}

	Model comparison: GRNN is best, but TXL complementary
\end{frame}

\begin{frame}{Why such large effects?}
	\pause
	Framework: Look at words long enough to ID with some threshold of certainty \pause

Could hypothesize mechanisms for difference: \pause
\begin{itemize}
	\item Task leads to higher threshold \pause 
	\item Task demands reduce available resources for processing \pause
	\item Presence of second word 
\end{itemize}

\end{frame}

\begin{frame}{Conclusion}

	\pause

Consider A-maze! 
\begin{itemize}\pause
	\item Documentation:  \textcolor{ForestGreen}{vboyce.github.io/Maze} \pause
	\item Versatile \pause
	\item Low spillover \pause 
\end{itemize}

Natural Stories A-maze: \pause 
\begin{itemize}
	\item Participants comprehend what they read \pause
	\item Find linear, large surprisal effects
\end{itemize}

\end{frame}

\appendix

\begin{frame}{Surprisal Effects}
	GAM if we only exclude mistakes (all participants, post-mistake data)
	\includegraphics[width=.9\textwidth]{../gam-any.pdf}	
\end{frame}

\end{document}