### 2021-07-05 Analysis/models.Rmd
more gam models

---
title: "Natural Stories model analyses"
output: html_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
library(mgcv)
library(mgcViz)
library(tidymv)
library(rsample)
library(cowplot)
library(scales)
library(here)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

```{r participants}

data <- read_rds(here("Data/cleaned.rds"))

data_filt <- data %>% filter(native %in% c("ENG", "English", "ENGLISH", "english")) #I peeked at what people put that semantically maps to english

data_error_summ <- data_filt %>% 
  mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
  group_by(subject) %>%
  filter(type!="practice") %>% 
  filter(rt<5000) %>% 
  summarize(pct_correct=mean(correct.num)) %>% 
  ungroup() %>% 
  mutate(is.attentive=ifelse(pct_correct>.8, T,F)) %>% 
  select(subject, is.attentive)

data_low_error <- data_filt %>% 
  left_join(data_error_summ, by="subject") %>% 
  filter(is.attentive) %>% 
  filter(type!="practice")

  data_error_free <- data_low_error %>% 
    mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
    group_by(sentence, subject) %>% fill(word_num_mistake) %>% ungroup() %>% 
    mutate(after_mistake=word_num-word_num_mistake,
           after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
    filter(correct=="yes") %>% 
    filter(!after_mistake %in% c(1,2))

data_no_first <- data_error_free %>% filter(word_num!=0)

data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  select(subject, word_num, word, rt, sentence, type)

data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  filter(after_mistake==0) %>% 
  select(subject, word_num, word, rt, sentence, type)

data_stories <- data_ready %>% select(type, subject) %>% 
  unique() %>% 
  group_by(type) %>% 
  tally()
```

<!--# Overview

`r data %>% select(subject) %>% unique() %>% nrow()` participants read naturalistic stories from the natural stories corpus. Each participant read 1 story.

We exclude

- participants who do not report English as a native language (`r data_filt %>% select(subject) %>% unique() %>% nrow()` remaining)
- participants who do not get 80% of the words correct (`r data_low_error %>% select(subject) %>% unique() %>% nrow()` remaining)
- practice items (`r data_low_error %>% nrow()` words remaining)
- words that were wrong or were within two after a mistake (`r data_error_free %>% nrow()` words remaining)
- the first word of every sentence (didn't have a real distractor, RT is measured slightly differently) (`r data_no_first %>% nrow()` words remaining)
- words with RTs <100 or >5000 (<100 we think is likely a recording error, or at least not reading the words at all, >5000 is likely getting distracted) (`r data_ready %>% nrow()` words remaining)

Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times, for an average of `r mean(data_stories$n)`.

We also do the analyses on only the words before mistakes (per sentence) (`r data_pre_error %>% nrow()` words)

From the modelling side:
(After attempts without doing this filtering) we only include words which are single token and known words in each of the models vocabularies. We also only include words with frequencies. This is roughly equivalent to excluding words with punctuation.

We use as predictors:

- length in characters of stripped word
- unigram frequency of word. Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances. (This tends to tokenize off punctuation, but is capitalization sensitive). Frequencies are represented as log2 of the expected occurances in 1 billion words.

Surprisals are measured in bits.

- ngram (5-gram KN smoothed)
- GRNN
- Transformer-XL

For GAM models, we center length and frequency but not surprisal. We want surprisal interpretable, but we also will be plotting it (at least for the bootstrapping) at length and frequencies set to 0, so they need to be centered. (Not sure this last piece is actually true/matters). -->

# Maze Prep
```{r labels}
labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())

select_labs <- labs %>% 
  mutate(across(ends_with("surp"),~ifelse(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1 & gpt_token_count==1 & word_num>0, .x, NA))) %>% 
  select(-ends_with("token_count")) %>% 
    mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
         freq_center=freq-mean(freq, na.rm=T),
         length_center=length-mean(length, na.rm=T),
         gpt_center=gpt_surp-mean(gpt_surp, na.rm=T)) %>% 
  group_by(sentence,Story_Num,Sentence_Num) %>% 
  mutate(across(txl_surp:gpt_center,.names="past_{.col}", lag),
         across(txl_surp:gpt_center,.names="past2_{.col}", ~lag(.x,n=2)),
         across(txl_surp:gpt_center,.names="past3_{.col}", ~lag(.x,n=3)))

```

```{r maze}

labelled_pre_error <- data_pre_error %>% inner_join(select_labs, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_")))%>% write_rds(here("Data/maze_pre_error.rds"))

```

```{r item-means}

select_data <- read_rds(here("Data/maze_pre_error.rds")) %>%  select(-starts_with("past2"), -starts_with("past3")) %>% 
  filter(across(everything(),~!is.na(.x)))

item_mean_data <- select_data %>% 
  group_by(across(c(-subject, -rt))) %>% 
  summarize(rt=mean(rt)) %>% 
  ungroup()

```

# Maze GAMs

```{r for gam}

ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                            ) %>% mutate(model="5-gram") 
grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                            )%>% mutate(model="GRNN")

txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                            ) %>% mutate(model="TXL")

gpt_data <- labelled_pre_error %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                            ) %>% mutate(model="GPT-2")

ngram_mean_data <- item_mean_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            ) %>% mutate(model="5-gram")

grnn_mean_data <- item_mean_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            )%>% mutate(model="GRNN")

txl_mean_data <- item_mean_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            ) %>% mutate(model="TXL")

gpt_mean_data <- item_mean_data %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            ) %>% mutate(model="GPT-2")

all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
  select(surprisal, prev_surp,model) %>% 
  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
  mutate(s=ifelse(name=="surprisal", "Current","Previous"))

all_mean_data <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data) %>% 
  select(surprisal, prev_surp,model) %>% 
  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
  mutate(s=ifelse(name=="surprisal", "Current","Previous")) %>% write_rds(here("Analysis/models/meaned_maze.rds"))
```

```{r formulae, echo=T}
#no hierarchical, has freq x len interaction
formula_interact <- rt ~ s(surprisal, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     ti(freq,len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")+
                     ti(prev_freq, prev_len, bs="cr")

# no hierarchical, NO freq x len interaction
formula_no_interact <- rt ~ s(surprisal, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")

formula_heirarchy <- rt ~ s(surprisal, bs="cr", k=20)+
                    s(surprisal, by=subject, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                      s(prev_surp, by=subject, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")

```

All of this is on by-item mean data.

```{r meaned data, eval=F}
gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
gam_grnn_1 <- gam(formula_interact, data=grnn_mean_data, method="REML")
gam_txl_1 <- gam(formula_interact, data=txl_mean_data, method="REML")
gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")

gam_ngram_2 <- gam(formula_no_interact, data=ngram_mean_data, method="REML")
gam_grnn_2 <- gam(formula_no_interact, data=grnn_mean_data, method="REML")
gam_txl_2 <- gam(formula_no_interact, data=txl_mean_data, method="REML")
gam_gpt_2 <- gam(formula_no_interact, data=gpt_mean_data, method="REML")

```

```{r anovas, eval=F}

anova(gam_ngram_1,gam_ngram_2)
anova(gam_grnn_1,gam_grnn_2)
anova(gam_txl_1,gam_txl_2)
anova(gam_gpt_1,gam_gpt_2)
```

I don't know how to interpret the above.

I also don't know how to interpret the plots.

```{r plots, eval=F}
plot(gam_gpt_1)
```

By-factor GAMs fail to run overnight / crash the R session.

```{r plot-gam, eval=F}

a <- get_gam_predictions(model=gam_ngram_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
b <- get_gam_predictions(model=gam_ngram_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
c <- get_gam_predictions(model=gam_grnn_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
d <- get_gam_predictions(model=gam_grnn_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
e <- get_gam_predictions(model=gam_txl_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
g <- get_gam_predictions(model=gam_gpt_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Current")
h <- get_gam_predictions(model=gam_gpt_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Previous")
all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% union(g) %>% union(h) %>% 
  mutate(model=factor(model, levels=c("5-gram", "GRNN", "TXL","GPT-2"))) %>% 
                        write_rds(here("Analysis/models/gam_predictions.rds"))
```

```{r, eval=F}

gam1 <- read_rds(here("Analysis/models/gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
  #facet_grid(s~model, scales="free_y")+
  facet_grid(s~model)+
  coord_cartesian(xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))

dens2 <-   ggplot(all_mean_data, aes(x=value))+
  geom_density(fill="gray",)+
  facet_grid(.~model)+
  labs(x="Surprisal (bits)", y="")+
    coord_cartesian(xlim=c(0,28))+
  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))

p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))

#confirm that this really is just a prettied up version of what the usual output is
#plot(gam_txl_1, seWithMean = TRUE, shift = coef(gam_txl_1)[1], select=1)

```

# SPR Prep
```{r read_data, eval=F}
d.raw.a <- here("Data/SPR/batch1_pro.csv") %>% 
  read_csv()

d.raw.b <- here("Data/SPR/batch2_pro.csv") %>% 
  read_csv()

offset=230 # following note in repo, we have to correct alignment on item 3 b/c an empty token was displayed

d.raw <- d.raw.a %>% union(d.raw.b) %>% 
  mutate(zone=if_else(item == 3 & zone > offset, zone - 3, zone - 2))

# spot check that this matches number wise with their processed words

```

```{r first-story, eval=F}
# to make things more comparable, only look at the first story each participant read

first <- d.raw %>% select(WorkerId,correct,item) %>% unique() %>% 
  group_by(WorkerId,correct) %>% 
  mutate(num=row_number()) %>% 
  filter(num==1) %>% 
  select(-num)

d.limited <- d.raw %>% inner_join(first, by=c("WorkerId","correct","item"))

#d.limited%>% select(WorkerId,correct,item) %>% unique() %>% group_by(correct) %>% tally()  %>% summarize(pct=n/sum(n),correct=correct)

d.good <- d.limited %>% filter(correct>4) %>% rename(rt=RT) %>% filter(rt>100 & rt<5000)
```

```{r surprisals, eval=F}

tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
  read_delim(delim="\t") %>% 
  mutate(Word_In_Story_Num=zone,
         Story_Num=item)

#something is going on is story 2, sentence 4 His brother had blatantly peeked..." 
#they have it as "peaked", not sure what was displayed, but we can inner join
spr_labelled <- select_labs %>% inner_join(tokenization, by=c("word","Word_In_Story_Num","Story_Num")) %>% 
  left_join(d.good, by=c("item","zone")) %>% 
  rename(subject=WorkerId) %>% 
  select(-WorkTimeInSeconds) %>% 
  filter(word_num>1) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_")))%>% write_rds(here("Data/clean_spr.rds"))
```

```{r spr item-means}

select_spr <- read_rds(here("Data/clean_spr.rds")) %>% 
  filter(across(everything(),~!is.na(.x))) %>% 
  ungroup()

item_mean_spr <- select_spr %>% 
  group_by(across(c(-subject, -rt))) %>% 
  summarize(rt=mean(rt)) %>% 
  ungroup()

```

# SPR GAMs

```{r for spr gam, eval=F}

ngram_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "ngram_", ""))%>%
  select(rt, surp, freq_center, length_center,
         past_surp, past_freq_center, past_length_center,
         past2_surp, past2_freq_center, past2_length_center,
         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="5-gram")

grnn_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "grnn_", ""))%>%
  select(rt, surp, freq_center, length_center,
         past_surp, past_freq_center, past_length_center,
         past2_surp, past2_freq_center, past2_length_center,
         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="GRNN")

txl_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "txl_", ""))%>%
  select(rt, surp, freq_center, length_center,
         past_surp, past_freq_center, past_length_center,
         past2_surp, past2_freq_center, past2_length_center,
         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="TXL")

gpt_mean_spr<-item_mean_spr %>% rename_with(~str_replace_all(.x, "gpt_", ""))%>%
  select(rt, surp, freq_center, length_center,
         past_surp, past_freq_center, past_length_center,
         past2_surp, past2_freq_center, past2_length_center,
         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="GPT-2")

all_mean_spr <- ngram_mean_spr %>% union(grnn_mean_spr) %>% union(txl_mean_spr) %>% union(gpt_mean_spr) %>% 
  select(surp, past_surp, past2_surp, past3_surp,model) %>% 
  pivot_longer(cols=`surp`:`past3_surp`) %>% 
  mutate(s=ifelse(name=="surprisal", "Current","Previous")) %>% write_rds(here("Analysis/models/meaned_spr.rds"))
```

```{r formulae, echo=T}
#no hierarchical, has freq x len interaction
spr_formula_interact <- rt ~ s(surp, bs="cr", k=20)+
                     ti(length_center, bs="cr")+
                     ti(freq_center,length_center, bs="cr")+
                     s(past_surp, bs="cr", k=20)+
                     ti(past_freq_center, bs="cr")+
                     ti(past_length_center, bs="cr")+
                     ti(past_freq_center, past_length_center, bs="cr")+
                     s(past2_surp, bs="cr", k=20)+
                     ti(past2_freq_center, bs="cr")+
                     ti(past2_length_center, bs="cr")+
                     ti(past2_freq_center, past2_length_center, bs="cr")+
                     s(past3_surp, bs="cr", k=20)+
                     ti(past3_freq_center, bs="cr")+
                     ti(past3_length_center, bs="cr")+
                     ti(past3_freq_center, past3_length_center, bs="cr")

```

All of this is on by-item mean data.

```{r spr meaned data, eval=F}
gam_ngram_1 <- gam(spr_formula_interact, data=ngram_mean_spr, method="REML")
gam_grnn_1 <- gam(spr_formula_interact, data=grnn_mean_spr, method="REML")
gam_txl_1 <- gam(spr_formula_interact, data=txl_mean_spr, method="REML")
gam_gpt_1 <- gam(spr_formula_interact, data=gpt_mean_spr, method="REML")

```

```{r plot-gam-spr, eval=F}

a <- get_gam_predictions(model=gam_ngram_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
b <- get_gam_predictions(model=gam_ngram_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
c <- get_gam_predictions(model=gam_ngram_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="2Previous")
d <- get_gam_predictions(model=gam_ngram_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="3Previous")

e <- get_gam_predictions(model=gam_grnn_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
f <- get_gam_predictions(model=gam_grnn_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
g <- get_gam_predictions(model=gam_grnn_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="2Previous")
h <- get_gam_predictions(model=gam_grnn_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="3Previous")

i <- get_gam_predictions(model=gam_txl_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
j <- get_gam_predictions(model=gam_txl_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
k <- get_gam_predictions(model=gam_txl_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="2Previous")
l <- get_gam_predictions(model=gam_txl_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="3Previous")

m <- get_gam_predictions(model=gam_gpt_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Current")
n <- get_gam_predictions(model=gam_gpt_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Previous")
o <- get_gam_predictions(model=gam_gpt_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="2Previous")
p <- get_gam_predictions(model=gam_gpt_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="3Previous")

all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% union(g) %>% union(h) %>% 
   union(i) %>% union(j) %>% union(k) %>% union(l) %>% union(m) %>% union(n) %>% union(o) %>% union(p) %>% 
  mutate(model=factor(model, levels=c("5-gram", "GRNN", "TXL","GPT-2")),
         s=factor(s, levels=c("Current", "Previous", "2Previous", "3Previous"))) %>% 
                        write_rds(here("Analysis/models/spr_gam_predictions.rds"))
```

```{r, eval=T}

gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
  facet_grid(s~model)+
  coord_cartesian(xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))

dens2 <-   ggplot(all_mean_spr, aes(x=value))+
  geom_density(fill="gray",)+
  facet_grid(.~model)+
  labs(x="Surprisal (bits)", y="")+
    coord_cartesian(xlim=c(0,28))+
  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))

p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))

```

# Maze LMs

```{r for brms}

ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_center, prev_surp=past_ngram_center, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
                                            ) %>% mutate(model="5-gram") 
grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_center, prev_surp=past_grnn_center, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
                                            )%>% mutate(model="GRNN")

txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_center, prev_surp=past_txl_center, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
                                            ) %>% mutate(model="TXL")

gpt_data <- labelled_pre_error %>% select(rt, surprisal=gpt_center, prev_surp=past_gpt_center, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
                                            ) %>% mutate(model="GPT-2")

all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
  write_rds(here("Analysis/models/maze_pre_brms.rds"))

```

```{r run brms, eval=F}
# Note this was actually run on a cluster with a separate script that just had this
data=readRDS("maze_pre_brms.rds")
 priors <- c(
      set_prior("normal(1000, 1000)", class="Intercept"),
      set_prior("normal(0, 500)", class="b"),
      set_prior("normal(0, 500)", class="sd"),
      set_prior("lkj(1)",       class="cor"))
run_model <- function(model_spec,data,file_name){
  d <- data %>% filter(model==model_spec)
  m1 <- brm(rt ~ surprisal * len + freq * len +
              prev_surp * prev_len + prev_freq * prev_len+
              (surprisal * len + freq * len +
              prev_surp * prev_len + prev_freq * prev_len|subject)+(1|Word_ID),
            data=d,
            file=file_name,
            prior=priors,
            control=list(adapt_delta=.95))
}

run_model("5-gram",data,"5-gram_maze")
run_model("GRNN",data,"grnn_maze")
run_model("TXL",data,"txl_maze")
run_model("GPT-2",data,"gpt_maze")

```

```{r summary}

show_summary <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()

  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
    group_by(.variable) %>% 
    summarize(pct_above_0=mean(above_0)) %>% 
    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
    left_join(intervals, by=".variable") %>% 
    mutate(lower=round(.lower, digits=1),
           upper=round(.upper, digits=1),
           E=round(.value, digits=1),
           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
           Term=str_sub(.variable, 3, -1),
           ) %>% 
    select(Term, `Estimate`)

  stats
}
```

```{r process-maze-models, eval=F}
ngram <- read_rds(here("Analysis/models/5-gram_maze.rds")) %>% show_summary() %>% mutate(model="5-gram")
txl <- read_rds(here("Analysis/models/txl_maze.rds")) %>% show_summary() %>% mutate(model="TXL")
grnn <- read_rds(here("Analysis/models/grnn_maze.rds")) %>% show_summary() %>% mutate(model="GRNN")
gpt <- read_rds(here("Analysis/models/gpt_maze.rds")) %>% show_summary() %>% mutate(model="GPT-2")

summ <- ngram %>% union(txl) %>% union(grnn) %>% union(gpt) %>% write_rds(here("Analysis/models/brms_maze_summ.rds"))

```

# SPR LMs

```{r spr brms}

ngram_data <- select_spr %>% rename_with(~str_replace_all(.x, "ngram", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="5-gram")

grnn_data <- select_spr %>% rename_with(~str_replace_all(.x, "grnn", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID)%>% mutate(model="GRNN")

txl_data <- select_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="TXL")

gpt_data <-select_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="GPT-2")

all_spr <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
  write_rds(here("Analysis/models/spr_pre_brms.rds"))

ngram_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "ngram", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="5-gram")

grnn_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "grnn", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, Word_ID)%>% mutate(model="GRNN")

txl_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="TXL")

gpt_mean_data <-item_mean_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="GPT-2")

all_mean_spr <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data)

```

```{r spr lms-mean, eval=T}

run_lm <- function(data,modelspec){
  d <- data %>% filter(model==modelspec)
model_1 <- lmer(rt ~ surp_center * length_center + length_center * freq_center +
                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center +(1|Word_ID), data=d)
return(model_1)
}

m_ngram <- run_lm(all_mean_spr,"5-gram")
m_grnn <- run_lm(all_mean_spr, "GRNN")
m_txl <- run_lm(all_mean_spr,"TXL")
m_gpt <- run_lm(all_mean_spr,"GPT-2")

```

```{r lmsumm}
summary(m_ngram)
summary(m_grnn)
summary(m_txl)
summary(m_gpt)
```

```{r, eval=F}

data=readRDS("spr_pre_brms.rds")
 priors <- c(
      set_prior("normal(300, 100)", class="Intercept"),
      set_prior("normal(0, 5)", class="b"),
      set_prior("normal(0, 5)", class="sd"),
      set_prior("lkj(1)",       class="cor"))
run_model <- function(model_spec,data,file_name){
  d <- data %>% filter(model==model_spec)
  m1 <- brm(rt ~ surp_center * length_center + length_center * freq_center +
                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
              (surp_center * length_center + length_center * freq_center +
                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center |subject)+(1|Word_ID),
            data=d,
            file=file_name,
            prior=priors,
            control=list(adapt_delta=.95))
}

run_model("5-gram",data,"5-gram_spr")
run_model("GRNN",data,"grnn_spr")
run_model("TXL",data,"txl_spr")
run_model("GPT-2",data,"gpt_spr")

```