### 2021-07-14 Analysis/models.Rmd
work on nested model comparisons

---
title: "Natural Stories model analyses"
output: html_document
---

# Prep

```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
options(knitr.table.format = "html")
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
library(mgcv)
library(mgcViz)
library(tidymv)
library(rsample)
library(cowplot)
library(scales)
library(here)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

```{r participants}

data <- read_rds(here("Data/cleaned.rds"))

data_filt <- data %>% filter(native %in% c("ENG", "English", "ENGLISH", "english")) #I peeked at what people put that semantically maps to english

data_error_summ <- data_filt %>% 
  mutate(correct.num=ifelse(correct=="yes", 1,0)) %>% 
  group_by(subject) %>%
  filter(type!="practice") %>% 
  filter(rt<5000) %>% 
  summarize(pct_correct=mean(correct.num)) %>% 
  ungroup() %>% 
  mutate(is.attentive=ifelse(pct_correct>.8, T,F)) %>% 
  select(subject, is.attentive)

data_low_error <- data_filt %>% 
  left_join(data_error_summ, by="subject") %>% 
  filter(is.attentive) %>% 
  filter(type!="practice")

  data_error_free <- data_low_error %>% 
    mutate(word_num_mistake=ifelse(correct=="no", word_num,NA)) %>% 
    group_by(sentence, subject) %>% fill(word_num_mistake) %>% ungroup() %>% 
    mutate(after_mistake=word_num-word_num_mistake,
           after_mistake=ifelse(is.na(after_mistake),0,after_mistake)) %>% 
    filter(correct=="yes") %>% 
    filter(!after_mistake %in% c(1,2))

data_no_first <- data_error_free %>% filter(word_num!=0)

data_ready <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  select(subject, word_num, word, rt, sentence, type)

data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  filter(after_mistake==0) %>% 
  select(subject, word_num, word, rt, sentence, type)

data_stories <- data_ready %>% select(type, subject) %>% 
  unique() %>% 
  group_by(type) %>% 
  tally()
```

<!--# Overview

`r data %>% select(subject) %>% unique() %>% nrow()` participants read naturalistic stories from the natural stories corpus. Each participant read 1 story.

We exclude

- participants who do not report English as a native language (`r data_filt %>% select(subject) %>% unique() %>% nrow()` remaining)
- participants who do not get 80% of the words correct (`r data_low_error %>% select(subject) %>% unique() %>% nrow()` remaining)
- practice items (`r data_low_error %>% nrow()` words remaining)
- words that were wrong or were within two after a mistake (`r data_error_free %>% nrow()` words remaining)
- the first word of every sentence (didn't have a real distractor, RT is measured slightly differently) (`r data_no_first %>% nrow()` words remaining)
- words with RTs <100 or >5000 (<100 we think is likely a recording error, or at least not reading the words at all, >5000 is likely getting distracted) (`r data_ready %>% nrow()` words remaining)

Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times, for an average of `r mean(data_stories$n)`.

We also do the analyses on only the words before mistakes (per sentence) (`r data_pre_error %>% nrow()` words)

From the modelling side:
(After attempts without doing this filtering) we only include words which are single token and known words in each of the models vocabularies. We also only include words with frequencies. This is roughly equivalent to excluding words with punctuation.

We use as predictors:

- length in characters of stripped word
- unigram frequency of word. Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances. (This tends to tokenize off punctuation, but is capitalization sensitive). Frequencies are represented as log2 of the expected occurances in 1 billion words.

Surprisals are measured in bits.

- ngram (5-gram KN smoothed)
- GRNN
- Transformer-XL

For GAM models, we center length and frequency but not surprisal. We want surprisal interpretable, but we also will be plotting it (at least for the bootstrapping) at length and frequencies set to 0, so they need to be centered. (Not sure this last piece is actually true/matters). -->

# Maze Prep
```{r labels}
labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())

select_labs <- labs %>% 
  mutate(across(ends_with("surp"),~ifelse(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1 & gpt_token_count==1 & word_num>0, .x, NA))) %>% 
  select(-ends_with("token_count")) %>% 
    mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
         freq_center=freq-mean(freq, na.rm=T),
         length_center=length-mean(length, na.rm=T),
         gpt_center=gpt_surp-mean(gpt_surp, na.rm=T)) %>% 
  group_by(sentence,Story_Num,Sentence_Num) %>% 
  mutate(across(txl_surp:gpt_center,.names="past_{.col}", lag),
         across(txl_surp:gpt_center,.names="past2_{.col}", ~lag(.x,n=2)),
         across(txl_surp:gpt_center,.names="past3_{.col}", ~lag(.x,n=3)))

```

```{r maze}

labelled_pre_error <- data_pre_error %>% inner_join(select_labs, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_")))%>% write_rds(here("Data/maze_pre_error.rds"))

```

```{r item-means}

select_data <- read_rds(here("Data/maze_pre_error.rds")) %>%  select(-starts_with("past2"), -starts_with("past3")) %>% 
  filter(across(everything(),~!is.na(.x)))

item_mean_data <- select_data %>% 
  group_by(across(c(-subject, -rt))) %>% 
  summarize(rt=mean(rt)) %>% 
  ungroup()

```

# Maze GAMs

```{r for gam}

ngram_data <- select_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                            ) %>% mutate(model="5-gram") 
grnn_data <- select_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                            )%>% mutate(model="GRNN")

txl_data <- select_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                            ) %>% mutate(model="TXL")

gpt_data <- select_data %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                            ) %>% mutate(model="GPT-2")

ngram_mean_data <- item_mean_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            ) %>% mutate(model="5-gram")

grnn_mean_data <- item_mean_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            )%>% mutate(model="GRNN")

txl_mean_data <- item_mean_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            ) %>% mutate(model="TXL")

gpt_mean_data <- item_mean_data %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            ) %>% mutate(model="GPT-2")

all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
  select(surprisal, prev_surp,model) %>% 
  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
  mutate(s=ifelse(name=="surprisal", "Current","Previous"))

all_mean_data <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data) %>% 
  select(surprisal, prev_surp,model) %>% 
  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
  mutate(s=ifelse(name=="surprisal", "Current","Previous")) %>% write_rds(here("Analysis/models/meaned_maze.rds"))
```

```{r formulae, echo=T}
#no hierarchical, has freq x len interaction
formula_interact <- rt ~ s(surprisal, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     ti(freq,len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")+
                     ti(prev_freq, prev_len, bs="cr")

# no hierarchical, NO freq x len interaction
formula_no_interact <- rt ~ s(surprisal, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")

formula_heirarchy <- rt ~ s(surprisal, bs="cr", k=20)+
                    s(surprisal, by=subject, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                      s(prev_surp, by=subject, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")

```

All of this is on by-item mean data.

```{r meaned data, eval=T, cache=T}
gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
gam_grnn_1 <- gam(formula_interact, data=grnn_mean_data, method="REML")
gam_txl_1 <- gam(formula_interact, data=txl_mean_data, method="REML")
gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")

gam_ngram_2 <- gam(formula_no_interact, data=ngram_mean_data, method="REML")
gam_grnn_2 <- gam(formula_no_interact, data=grnn_mean_data, method="REML")
gam_txl_2 <- gam(formula_no_interact, data=txl_mean_data, method="REML")
gam_gpt_2 <- gam(formula_no_interact, data=gpt_mean_data, method="REML")

```

```{r anovas, eval=T}

anova(gam_ngram_1,gam_ngram_2, test="Chisq")
anova(gam_grnn_1,gam_grnn_2, test="Chisq")
anova(gam_txl_1,gam_txl_2, test="Chisq")
anova(gam_gpt_1,gam_gpt_2, test="Chisq")
```

```{r plots, eval=T}
plot(gam_gpt_1, select=4)
plot(gam_grnn_1, select=4)

plot(gam_txl_1, select=4)
plot(gam_ngram_1, select=4)

```

It looks like the effect is ~ 0 for most of the space where there's actually data, and it only goes off the rails in the region where there isn't data.

Below is all words with centered frequency < -6 and centered_length > 4. Most of these are I think are recognizable as words. 
```{r}
select_data %>% select(word,  freq_center, length_center, sentence) %>% unique() %>% 
  arrange(freq_center) %>% 
  filter(freq_center< -6 & length_center > 4) %>% knitr::kable()
```

```{r plot-gam, eval=F}

a <- get_gam_predictions(model=gam_ngram_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
b <- get_gam_predictions(model=gam_ngram_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
c <- get_gam_predictions(model=gam_grnn_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
d <- get_gam_predictions(model=gam_grnn_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
e <- get_gam_predictions(model=gam_txl_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
f <- get_gam_predictions(model=gam_txl_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
g <- get_gam_predictions(model=gam_gpt_1, series=surprisal, series_length=100) %>% select(surprisal, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Current")
h <- get_gam_predictions(model=gam_gpt_1, series=prev_surp, series_length=100) %>% select(surprisal=prev_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Previous")
all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% union(g) %>% union(h) %>% 
  mutate(model=factor(model, levels=c("5-gram", "GRNN", "TXL","GPT-2"))) %>% 
                        write_rds(here("Analysis/models/gam_predictions.rds"))
```

```{r, eval=T}

gam1 <- read_rds(here("Analysis/models/gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
  #facet_grid(s~model, scales="free_y")+
  facet_grid(s~model)+
  coord_cartesian(xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))

dens2 <-   ggplot(all_mean_data, aes(x=value))+
  geom_density(fill="gray",)+
  facet_grid(.~model)+
  labs(x="Surprisal (bits)", y="")+
    coord_cartesian(xlim=c(0,28))+
  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))

p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))

#confirm that this really is just a prettied up version of what the usual output is
#plot(gam_txl_1, seWithMean = TRUE, shift = coef(gam_txl_1)[1], select=1)

```

# Maze LMs

```{r for brms}

ngram_data <- labelled_pre_error %>% select(rt, surprisal=ngram_center, prev_surp=past_ngram_center, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
                                            ) %>% mutate(model="5-gram") 
grnn_data <- labelled_pre_error %>% select(rt, surprisal=grnn_center, prev_surp=past_grnn_center, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
                                            )%>% mutate(model="GRNN")

txl_data <- labelled_pre_error %>% select(rt, surprisal=txl_center, prev_surp=past_txl_center, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
                                            ) %>% mutate(model="TXL")

gpt_data <- labelled_pre_error %>% select(rt, surprisal=gpt_center, prev_surp=past_gpt_center, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject, Word_ID
                                            ) %>% mutate(model="GPT-2")

all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
  write_rds(here("Analysis/models/maze_pre_brms.rds"))

```

```{r run brms, eval=F}
# Note this was actually run on a cluster with a separate script that just had this
data=readRDS("maze_pre_brms.rds")
 priors <- c(
      set_prior("normal(1000, 1000)", class="Intercept"),
      set_prior("normal(0, 500)", class="b"),
      set_prior("normal(0, 500)", class="sd"),
      set_prior("lkj(1)",       class="cor"))
run_model <- function(model_spec,data,file_name){
  d <- data %>% filter(model==model_spec)
  m1 <- brm(rt ~ surprisal * len + freq * len +
              prev_surp * prev_len + prev_freq * prev_len+
              (surprisal * len + freq * len +
              prev_surp * prev_len + prev_freq * prev_len|subject)+(1|Word_ID),
            data=d,
            file=file_name,
            prior=priors,
            control=list(adapt_delta=.95))
}

run_model("5-gram",data,"5-gram_maze")
run_model("GRNN",data,"grnn_maze")
run_model("TXL",data,"txl_maze")
run_model("GPT-2",data,"gpt_maze")

```

```{r summary}

show_summary <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()

  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
    group_by(.variable) %>% 
    summarize(pct_above_0=mean(above_0)) %>% 
    mutate(`P` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
    left_join(intervals, by=".variable") %>% 
    mutate(lower=round(.lower, digits=1),
           upper=round(.upper, digits=1),
           E=round(.value, digits=1),
           `Estimate`=str_c(E," [",lower,", ", upper,"]"),
           Term=str_sub(.variable, 3, -1),
           ) %>% 
    select(Term, `Estimate`)

  stats
}
```

```{r process-maze-models, eval=F}
ngram <- read_rds(here("Analysis/models/5-gram_maze.rds")) %>% show_summary() %>% mutate(model="5-gram")
txl <- read_rds(here("Analysis/models/txl_maze.rds")) %>% show_summary() %>% mutate(model="TXL")
grnn <- read_rds(here("Analysis/models/grnn_maze.rds")) %>% show_summary() %>% mutate(model="GRNN")
gpt <- read_rds(here("Analysis/models/gpt_maze.rds")) %>% show_summary() %>% mutate(model="GPT-2")

summ <- ngram %>% union(txl) %>% union(grnn) %>% union(gpt) %>% write_rds(here("Analysis/models/brms_maze_summ.rds"))

```

# Maze Model comparison

I'm not sure if this is what we want, but these are the P-values from anova's comparison the value of adding one set of surprisal predictors to a model that already has another. (So smaller values mean that the column  has a lot of predictive utility beyond the row label).

I'm not sure what the best presentational format is -- or whether we want F values or p values or what. I also not sure how to arrange the table (should I transpose?) or label it so which one is better is clear.

This is all on meaned data with no mixed effects and predictors of freq x length and surprisal(s) for word and previous word.

```{r}
select_data <- read_rds(here("Data/maze_pre_error.rds")) %>%  select(-starts_with("past2"), -starts_with("past3")) %>% 
  filter(across(everything(),~!is.na(.x)))

item_mean_data <- select_data %>% 
  group_by(across(c(-subject, -rt))) %>% 
  summarize(rt=mean(rt)) %>% 
  ungroup()

no_surp <- "rt ~ freq_center * length_center + past_freq_center * past_length_center"
ngram <- "+ ngram_center + past_ngram_center"
grnn <- "+ grnn_center + past_grnn_center"
txl <- "+ txl_center + past_txl_center"
gpt <- "+ gpt_center + past_gpt_center"

do_comparison <- function(m1, m2){
  # is there benefit of m2 above m1
  if (m1==m2) {return("")}
  lm1 <- lm(str_c(no_surp,m1), data=item_mean_data)
  lm2 <- lm(str_c(no_surp,m1,m2), data=item_mean_data)
  a <- anova(lm1,lm2)
  fval <- anova(lm1,lm2)$F[2] %>% round() %>%  as.character
  pval <- anova(lm1,lm2)$`Pr(>F)`[2] %>% pvalue(add_p=T) %>% as.character()
  return(str_c(fval," (",pval,")"))
}

log_like <- function(m1){
    lm(str_c(no_surp,m1), data=item_mean_data) %>% logLik() %>% round() %>% as.character()

}

tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
             `over GRNN`=grnn,`over TXL`=txl,`over GPT-2`=gpt) %>% 
   mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
          `Log Lik`=map_chr(model_str, log_like)) %>% 
   select(-model_str) %>% write_rds(here("Analysis/maze_model_compare.rds")) %>%  knitr::kable()

```

# SPR Prep
```{r read_data, eval=F}
d.raw.a <- here("Data/SPR/batch1_pro.csv") %>% 
  read_csv()

d.raw.b <- here("Data/SPR/batch2_pro.csv") %>% 
  read_csv()

offset=230 # following note in repo, we have to correct alignment on item 3 b/c an empty token was displayed

d.raw <- d.raw.a %>% union(d.raw.b) %>% 
  mutate(zone=if_else(item == 3 & zone > offset, zone - 3, zone - 2))

# spot check that this matches number wise with their processed words

```

```{r first-story, eval=F}
# to make things more comparable, only look at the first story each participant read

first <- d.raw %>% select(WorkerId,correct,item) %>% unique() %>% 
  group_by(WorkerId,correct) %>% 
  mutate(num=row_number()) %>% 
  filter(num==1) %>% 
  select(-num)

d.limited <- d.raw %>% inner_join(first, by=c("WorkerId","correct","item"))

#d.limited%>% select(WorkerId,correct,item) %>% unique() %>% group_by(correct) %>% tally()  %>% summarize(pct=n/sum(n),correct=correct)

d.good <- d.limited %>% filter(correct>4) %>% rename(rt=RT) %>% filter(rt>100 & rt<5000)
```

```{r surprisals, eval=F}

tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
  read_delim(delim="\t") %>% 
  mutate(Word_In_Story_Num=zone,
         Story_Num=item)

#something is going on is story 2, sentence 4 His brother had blatantly peeked..." 
#they have it as "peaked", not sure what was displayed, but we can inner join
spr_labelled <- select_labs %>% inner_join(tokenization, by=c("word","Word_In_Story_Num","Story_Num")) %>% 
  left_join(d.good, by=c("item","zone")) %>% 
  rename(subject=WorkerId) %>% 
  select(-WorkTimeInSeconds) %>% 
  filter(word_num>1) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_")))%>% write_rds(here("Data/clean_spr.rds"))
```

```{r spr item-means}

select_spr <- read_rds(here("Data/clean_spr.rds")) %>% 
  filter(across(everything(),~!is.na(.x))) %>% 
  ungroup()

item_mean_spr <- select_spr %>% 
  group_by(across(c(-subject, -rt))) %>% 
  summarize(rt=mean(rt)) %>% 
  ungroup()

```

# SPR GAMs

```{r for spr gam, eval=F}

ngram_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "ngram_", ""))%>%
  select(rt, surp, freq_center, length_center,
         past_surp, past_freq_center, past_length_center,
         past2_surp, past2_freq_center, past2_length_center,
         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="5-gram")

grnn_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "grnn_", ""))%>%
  select(rt, surp, freq_center, length_center,
         past_surp, past_freq_center, past_length_center,
         past2_surp, past2_freq_center, past2_length_center,
         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="GRNN")

txl_mean_spr <- item_mean_spr %>% rename_with(~str_replace_all(.x, "txl_", ""))%>%
  select(rt, surp, freq_center, length_center,
         past_surp, past_freq_center, past_length_center,
         past2_surp, past2_freq_center, past2_length_center,
         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="TXL")

gpt_mean_spr<-item_mean_spr %>% rename_with(~str_replace_all(.x, "gpt_", ""))%>%
  select(rt, surp, freq_center, length_center,
         past_surp, past_freq_center, past_length_center,
         past2_surp, past2_freq_center, past2_length_center,
         past3_surp, past3_freq_center, past3_length_center) %>% mutate(model="GPT-2")

all_mean_spr <- ngram_mean_spr %>% union(grnn_mean_spr) %>% union(txl_mean_spr) %>% union(gpt_mean_spr) %>% 
  select(surp, past_surp, past2_surp, past3_surp,model) %>% 
  pivot_longer(cols=`surp`:`past3_surp`) %>% 
  mutate(s=ifelse(name=="surprisal", "Current","Previous")) %>% write_rds(here("Analysis/models/meaned_spr.rds"))
```

```{r spr formulae, echo=T}
#no hierarchical, has freq x len interaction
spr_formula_interact <- rt ~ s(surp, bs="cr", k=20)+
                     ti(length_center, bs="cr")+
                     ti(freq_center,length_center, bs="cr")+
                     s(past_surp, bs="cr", k=20)+
                     ti(past_freq_center, bs="cr")+
                     ti(past_length_center, bs="cr")+
                     ti(past_freq_center, past_length_center, bs="cr")+
                     s(past2_surp, bs="cr", k=20)+
                     ti(past2_freq_center, bs="cr")+
                     ti(past2_length_center, bs="cr")+
                     ti(past2_freq_center, past2_length_center, bs="cr")+
                     s(past3_surp, bs="cr", k=20)+
                     ti(past3_freq_center, bs="cr")+
                     ti(past3_length_center, bs="cr")+
                     ti(past3_freq_center, past3_length_center, bs="cr")

```

All of this is on by-item mean data.

```{r spr meaned data, eval=F}
gam_ngram_1 <- gam(spr_formula_interact, data=ngram_mean_spr, method="REML")
gam_grnn_1 <- gam(spr_formula_interact, data=grnn_mean_spr, method="REML")
gam_txl_1 <- gam(spr_formula_interact, data=txl_mean_spr, method="REML")
gam_gpt_1 <- gam(spr_formula_interact, data=gpt_mean_spr, method="REML")

```

```{r plot-gam-spr, eval=F}

a <- get_gam_predictions(model=gam_ngram_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Current")
b <- get_gam_predictions(model=gam_ngram_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="Previous")
c <- get_gam_predictions(model=gam_ngram_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="2Previous")
d <- get_gam_predictions(model=gam_ngram_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="5-gram", s="3Previous")

e <- get_gam_predictions(model=gam_grnn_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Current")
f <- get_gam_predictions(model=gam_grnn_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="Previous")
g <- get_gam_predictions(model=gam_grnn_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="2Previous")
h <- get_gam_predictions(model=gam_grnn_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GRNN", s="3Previous")

i <- get_gam_predictions(model=gam_txl_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Current")
j <- get_gam_predictions(model=gam_txl_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="Previous")
k <- get_gam_predictions(model=gam_txl_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="2Previous")
l <- get_gam_predictions(model=gam_txl_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="TXL", s="3Previous")

m <- get_gam_predictions(model=gam_gpt_1, series=surp, series_length=100) %>% select(surprisal=surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Current")
n <- get_gam_predictions(model=gam_gpt_1, series=past_surp, series_length=100) %>% select(surprisal=past_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="Previous")
o <- get_gam_predictions(model=gam_gpt_1, series=past2_surp, series_length=100) %>% select(surprisal=past2_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="2Previous")
p <- get_gam_predictions(model=gam_gpt_1, series=past3_surp, series_length=100) %>% select(surprisal=past3_surp, rt, CI_upper, CI_lower) %>% unique() %>% mutate(model="GPT-2", s="3Previous")

all <- a %>% union(b) %>% union(c) %>% union(d) %>% union(e) %>% union(f) %>% union(g) %>% union(h) %>% 
   union(i) %>% union(j) %>% union(k) %>% union(l) %>% union(m) %>% union(n) %>% union(o) %>% union(p) %>% 
  mutate(model=factor(model, levels=c("5-gram", "GRNN", "TXL","GPT-2")),
         s=factor(s, levels=c("Current", "Previous", "2Previous", "3Previous"))) %>% 
                        write_rds(here("Analysis/models/spr_gam_predictions.rds"))
```

```{r, eval=T}
all_mean_spr <-  read_rds(here("Analysis/models/meaned_spr.rds"))

gam1 <- read_rds(here("Analysis/models/spr_gam_predictions.rds")) %>%  ggplot( aes(x=surprisal, y=rt, ymin=CI_lower, ymax=CI_upper))+
  geom_line()+
  geom_ribbon(alpha=.3)+
  facet_grid(s~model)+
  coord_cartesian(xlim=c(0,28))+
  labs(x="Surprisal (bits)", y="Reaction Time (ms)")+theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), plot.margin=margin(t=0,r=0,b=0,l=0,unit="pt"))

dens2 <-   ggplot(all_mean_spr, aes(x=value))+
  geom_density(fill="gray",)+
  facet_grid(.~model)+
  labs(x="Surprisal (bits)", y="")+
    coord_cartesian(xlim=c(0,28))+
  theme(axis.text.y = element_blank(), strip.text=element_blank(), axis.ticks.y =element_blank(), axis.title.y=element_blank(), panel.grid=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "cm"))

p2 = cowplot::align_plots(gam1, dens2, align = "v", axis="lr")
plot_grid(p2[[1]], p2[[2]], nrow=2, rel_heights = c(1, .3))

```

# SPR LMs

```{r spr brms, eval=F}

select_spr <- read_rds(here("Data/clean_spr.rds")) %>% 
  filter(across(everything(),~!is.na(.x))) %>% 
  ungroup()

item_mean_spr <- select_spr %>% 
  group_by(across(c(-subject, -rt))) %>% 
  summarize(rt=mean(rt)) %>% 
  ungroup()

ngram_data <- select_spr %>% rename_with(~str_replace_all(.x, "ngram", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="5-gram")

grnn_data <- select_spr %>% rename_with(~str_replace_all(.x, "grnn", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID)%>% mutate(model="GRNN")

txl_data <- select_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="TXL")

gpt_data <-select_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, subject, Word_ID) %>% mutate(model="GPT-2")

all_spr <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
  write_rds(here("Analysis/models/spr_pre_brms.rds"))

ngram_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "ngram", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="5-gram")

grnn_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "grnn", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, Word_ID)%>% mutate(model="GRNN")

txl_mean_data <- item_mean_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="TXL")

gpt_mean_data <-item_mean_spr %>% rename_with(~str_replace_all(.x, "txl", "surp"))%>% 
  select(rt, surp_center, freq_center, length_center,
         past_surp_center, past_freq_center, past_length_center,
         past2_surp_center, past2_freq_center, past2_length_center,
         past3_surp_center, past3_freq_center, past3_length_center, Word_ID) %>% mutate(model="GPT-2")

all_mean_spr_center <- ngram_mean_data %>% union(grnn_mean_data) %>% union(txl_mean_data) %>% union(gpt_mean_data) %>% write_rds(here("Analysis/models/mean_spr_center.rds"))

```

```{r spr lms-mean, eval=T}

all_mean_spr_center <- read_rds(here("Analysis/models/mean_spr_center.rds"))
run_lm <- function(data,modelspec){
  d <- data %>% filter(model==modelspec)
model_1 <- lmer(rt ~ surp_center * length_center + length_center * freq_center +
                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center +(1|Word_ID), data=d)
return(model_1)
}

m_ngram <- run_lm(all_mean_spr_center,"5-gram")
m_grnn <- run_lm(all_mean_spr_center, "GRNN")
m_txl <- run_lm(all_mean_spr_center,"TXL")
m_gpt <- run_lm(all_mean_spr_center,"GPT-2")

```

```{r lmsumm}
summary(m_ngram)
summary(m_grnn)
summary(m_txl)
summary(m_gpt)
```

```{r, eval=F}

#data=readRDS("spr_pre_brms.rds")
data=read_rds(here("Analysis/models/spr_pre_brms.rds"))
 priors <- c(
      set_prior("normal(0, 5)", class="b"),
      set_prior("exponential(.1)", class="sd", group="subject", coef="Intercept"),
      set_prior("exponential(.1)", class="sd", group="Word_ID", coef="Intercept"),
      set_prior("exponential(1)", class="sd"),
      set_prior("lkj(2)",       class="cor"),
      set_prior("exponential(.1)", class="sigma")
      )
run_model <- function(model_spec,data,file_name){
  d <- data %>% filter(model==model_spec) %>% 
    mutate(center_rt=scale(rt, center=T, scale=F))
  m1 <- brm(center_rt ~ surp_center * length_center + length_center * freq_center +
                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
              (surp_center + length_center + freq_center|subject)+(1|Word_ID),
            data=d,
            file=file_name,
            prior=priors,
            control=list(adapt_delta=.95))
}

#run_model("5-gram",data,"5-gram_spr")
#run_model("GRNN",data,"grnn_spr")
run_model("TXL",data,"txl_spr")
#run_model("GPT-2",data,"gpt_spr")

```

```{r diagnose-unhappiness}

sad <- read_rds(here("Analysis/models/gpt_spr.rds"))

summary(sad)

d <- data %>% filter(model=="GPT-2") %>% select(-rt, -Word_ID, -model, -subject)

cor(d) %>%as.tibble() %>%  mutate(across(where(is.numeric), ~round(.x,2))) %>% View()
```

Really struggling with the intercept for population & subject level -- maybe sd was intercept needs to be a lot bigger?

# SPR Model comparison

I'm not sure if this is what we want, but these are the P-values from anova's comparison the value of adding one set of surprisal predictors to a model that already has another. (So smaller values mean that the column  has a lot of predictive utility beyond the row label).

I'm not sure what the best presentational format is -- or whether we want F values or p values or what. I also not sure how to arrange the table (should I transpose?) or label it so which one is better is clear.

This is all on meaned data with no mixed effects and predictors of freq x length and surprisal(s) for word and previous word.

```{r}
select_spr <- read_rds(here("Data/clean_spr.rds")) %>% 
  filter(across(everything(),~!is.na(.x))) %>% 
  ungroup()

item_mean_spr <- select_spr %>% 
  group_by(across(c(-subject, -rt))) %>% 
  summarize(rt=mean(rt)) %>% 
  ungroup()

no_surp <- "rt ~ freq_center * length_center + 
            past_freq_center * past_length_center + 
            past2_freq_center * past2_length_center + 
            past3_freq_center * past3_length_center "
ngram <- "+ ngram_center + past_ngram_center + past2_ngram_center + past3_ngram_center"
grnn <- "+ grnn_center + past_grnn_center +past2_grnn_center +past3_grnn_center"
txl <- "+ txl_center + past_txl_center +past2_txl_center + past3_txl_center"
gpt <- "+ gpt_center + past_gpt_center + past2_gpt_center + + past3_gpt_center"

do_comparison <- function(m1, m2){
  # is there benefit of m2 above m1
  if (m1==m2) {return("")}
  lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr)
  lm2 <- lm(str_c(no_surp,m1,m2), data=item_mean_spr)
  a <- anova(lm1,lm2)
  fval <- anova(lm1,lm2)$F[2] %>% round() %>%  as.character
  pval <- anova(lm1,lm2)$`Pr(>F)`[2] %>% pvalue(add_p=T) %>% as.character()
  return(str_c(fval," (",pval,")"))
}

log_like <- function(m1){
    lm(str_c(no_surp,m1), data=item_mean_spr) %>% logLik() %>% round() %>% as.character()

}

tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
             `over GRNN`=grnn,`over TXL`=txl,`over GPT-2`=gpt) %>% 
   mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
          `Log Lik`=map_chr(model_str, log_like)) %>% 
   select(-model_str)%>% write_rds(here("Analysis/spr_model_compare.rds")) %>% kable()

```