### 2022-08-25 Papers/Cover_letter/glossapx_cover_1.tex
draft cover letter

\documentclass{stanfordletter}
%\makelabels
\begin{document}
	\name{Veronica Boyce}
	\signature{Veronica Boyce \\ Roger Levy}

	\begin{letter}{Fernanda Ferreira and Brian Dillon \\ Co-Editors-in-Chief \\ Glossa Psycholinguistics}

		\opening{Dear Professor Ferreira and Professor Dillon,} 
		 This submission contains the manuscript ``A-maze of Natural Stories: Comprehension and surprisal in the Maze task'' which we are submitting to Glossa Psycholinguistics for publication. This reports a variation on the Maze task designed to accommodate longer naturalistic texts and an experiment using this method where participants read passages from the Natural Stories corpus (Futrell et al 2020). We present an analysis of the reading time characteristics of the data, focusing on surprisal effects, and we compare to self-paced-reading data collected on the same materials by Futrell et al 2020.

		 Researchers with relevant expertise and interest who therefore might be appropriate reviewers include Jeffrey Witzel, Kenneth Forster, Viktor Kuperman, Dustin Chac{\'o}n, and Masaya Yoshida.

		 This manuscript is not under review at any other venue. Please let us know if you need any further information in
		 connection with this submission. We look forward to hearing from you regarding the manuscript!
		\closing{Sincerely,}

	\end{letter}

\end{document}