### 2022-08-25 Papers/Cover_letter/glossapx_cover_1.tex
Roger's edits

\documentclass{stanfordletter}
%\makelabels
\begin{document}
	\name{Veronica Boyce}
	\signature{Veronica Boyce \\ Roger Levy}

	\begin{letter}{Fernanda Ferreira and Brian Dillon \\ Co-Editors-in-Chief \\ Glossa Psycholinguistics}

          \opening{Dear Professor Ferreira and Professor Dillon,} This
          submission contains the manuscript ``A-maze of Natural
          Stories: Comprehension and surprisal in the Maze task''
          which we are submitting to Glossa Psycholinguistics for
          publication. The manuscript reports a variation on the Maze
          task (Forster et al., 2009; Boyce et al., 2020) designed to
          accommodate longer naturalistic texts and an experiment
          using this method where participants read passages from the
          Natural Stories corpus (Futrell et al 2020). We present an
          analysis of the reading time characteristics of the data,
          focusing on surprisal effects, and we compare to
          self-paced-reading data collected on the same materials by
          Futrell et al 2020. We find that in our new Maze task
          variant participants can successfully read and comprehend
          the longer Natural Stories corpus texts, that the task
          yields RTs that are better localized and more sensitive than
          self-paced reading RTs, and that the same linear effect of
          word surprisal on RTs previously documented in self-paced
          reading and eye tracking of naturalistic texts is obtained
          in our maze RT data. We believe this paper will be of broad
          methodological relevance, and also potential theoretical
          interest, for the psycholinguistics community.

          Researchers with relevant expertise and interest who
          therefore might be appropriate reviewers include Jeffrey
          Witzel, Kenneth Forster, Viktor Kuperman, Dustin Chac{\'o}n,
          and Masaya Yoshida.

          This manuscript is not under review at any other
          venue. Please let us know if you need any further
          information in connection with this submission. We look
          forward to hearing from you regarding the manuscript!
          \closing{Sincerely,}

	\end{letter}

\end{document}